### 2021-07-27 tex/lrsub.tex

With the notion of \emph{environment} decided, the proofs of admissibility of
renaming and substitution are structured in the same way as they are in
\autoref{sec:syntactic-kits}.
We start with a slightly modified definition of \AgdaRecord{Kit}.
While intuitionistic kits require arbitrary (right-)weakening, this assumption
is unrealistic in the linear setting.
However, we can expect weakening by $\gr0$-use variables.

\missingfigure{Kit}