### 2021-08-10 agda/LRKits.lagda.tex
[ new ] start on lR substitution

\begin{code}
{-# OPTIONS --safe --without-K --postfix-projections --prop #-}
open import Algebra.Po
open import Level

module LRKits (poSemiring : PoSemiring 0ℓ 0ℓ 0ℓ) where

  open PoSemiring poSemiring renaming (Carrier to Ann)

  open import Algebra.Relational renaming (_↘_↙_ to _↘-_-↙_)
  open import Data.LTree
  open import Data.LTree.Vector
  open import Data.Product
  open import Data.Wrap
  open import Function using (id; _∘_; Equivalence)
  open import Relation.Binary.PropositionalEquality as ≡ using (_≡_)
  open import Relation.Nary
  open import Relation.Unary.Bunched

  data Ty : Set where
    ι : Ty
    `! : (r : Ann) (A : Ty) → Ty

  open import Generic.Linear.Operations rawPoSemiring
  open import Generic.Linear.Algebra poSemiring
  open import Generic.Linear.Syntax Ty Ann
  open import Generic.Linear.Syntax.Interpretation Ty rawPoSemiring
  open import Generic.Linear.Variable Ty rawPoSemiring
  open import Generic.Linear.Environment Ty poSemiring
  open import Generic.Linear.Environment.Properties Ty poSemiring
  open import Generic.Linear.Environment.Categorical Ty poSemiring
\end{code}

%<*Tm>
\begin{code}
  Bind : Ctx → (Ctx → Set) → (Ctx → Set)
  Bind Δ T Γ = T (Γ ++ᶜ Δ)

  data _⊢_ : Ctx → Ty → Set where
    var : ∀[ _∋_ ⇒ _⊢_ ]
    !-I : ∀ {r A} → ∀[ r ·ᶜ (_⊢ A) ⇒ _⊢ `! r A ]
    !-E : ∀ {r A C} → ∀[ (_⊢ `! r A) ✴ᶜ Bind [ r · A ]ᶜ (_⊢ C) ⇒ _⊢ C ]
\end{code}
%</Tm>

\begin{code}
  module _ {𝓥 : Ctx → Ty → Set} where

    env-+ : ∀ {s t γ δ P Qₗ Qᵣ} →
      let Γ = ctx {s} P γ; Δₗ = ctx {t} Qₗ δ; Δᵣ = ctx {t} Qᵣ δ in
      (λ Q → [ 𝓥 ] Γ ⇒ᵉ ctx Q δ) ◇ _≤[ Qₗ +* Qᵣ ] →
      (λ Pₗ → [ 𝓥 ] ctx Pₗ γ ⇒ᵉ Δₗ)
        ↘- P ≤[_+*_] -↙
      (λ Pᵣ → [ 𝓥 ] ctx Pᵣ γ ⇒ᵉ Δᵣ)
    env-+ (ρ , sp) =
      let sumsₗ ↘, sp′ ,↙ sumsᵣ = ρ .asLinRel .linRel .rel-+ₘ (sp , ρ .sums) in
      record { [_]_⇒ᵉ_ ρ; sums = sumsₗ }
        ↘, sp′ ,↙
      record { [_]_⇒ᵉ_ ρ; sums = sumsᵣ }

    env-* : ∀ {s t γ δ P Q′ r} →
      let Γ = ctx {s} P γ; Δ′ = ctx {t} Q′ δ in
      (λ Q → [ 𝓥 ] Γ ⇒ᵉ ctx Q δ) ◇ _≤[ r *ₗ Q′ ] →
      P ≤[ r *ₗ_] ◇ (λ P′ → [ 𝓥 ] ctx P′ γ ⇒ᵉ Δ′)
    env-* (ρ , sp) =
      let sp′ , sums′ = ρ .asLinRel .linRel .rel-*ₘ (sp , ρ .sums) in
      sp′ , record { [_]_⇒ᵉ_ ρ; sums = sums′ }
\end{code}

%<*Kit>
\begin{code}
  Weakening : ∀ {v} {A : Set} → (Ctx → A → Set v) → Set v
  Weakening 𝓥 =
    ∀ {s t P} {γ : Vector Ty (s <+> t)} → P ∘ ↘ ≤* 0* →
    ∀[ 𝓥 (ctx (P ∘ ↙) (γ ∘ ↙)) ⇒ 𝓥 (ctx P γ) ]

  record Kit (𝓥 : Ctx → Ty → Set) : Set where
    constructor kit
    field
      ↙ᵏ : Weakening 𝓥
      vr : ∀[ _∋_ ⇒ 𝓥 ]
      tm : ∀[ 𝓥 ⇒ _⊢_ ]
\end{code}
%</Kit>

\begin{code}
    open Equivalence
\end{code}

%<*bindEnv>
\begin{code}
    bindEnv : ∀ {Γ Δ Θ} → [ 𝓥 ] Γ ⇒ᵉ Δ → [ 𝓥 ] Γ ++ᶜ Θ ⇒ᵉ Δ ++ᶜ Θ
    bindEnv ρ .M = [ [ ρ .M │ 0ᴹ ] ─ [ 0ᴹ │ 1ᴹ ] ]
    bindEnv ρ .asLinRel = ρ .asLinRel ⊕AsLinRel idAsLinRel
    bindEnv ρ .sums = ρ .sums , ≤*-refl
    bindEnv ρ .lookup (ll , rr) (lvar (↙ i) q b) =
      let bl , br = un++ₙ b in
      ↙ᵏ (≤*-trans rr br) (ρ .lookup ll (lvar i q bl))
    bindEnv ρ .lookup (ll , rr) (lvar (↘ i) q b) =
      let bl , br = un++ₙ b in
      let bl′ = 0*→≤* (ρ .asLinRel .linRel .rel-0ₘ (≤*→0* bl , ll)) in
      vr (lvar (↘ i) q (bl′ ++ₙ ≤*-trans rr br))
\end{code}
%</bindEnv>

%<*trav>
\begin{code}
    trav : ∀ {Γ Δ A} → [ 𝓥 ] Γ ⇒ᵉ Δ → Δ ⊢ A → Γ ⊢ A
    trav ρ (var x) = tm (ρ .lookup (ρ .sums) x)
    trav ρ (!-I (⟨ sp* ⟩· M)) =
      let sp*′ , ρ′ = env-* (ρ , sp*) in
      !-I (⟨ sp*′ ⟩· (trav ρ′ M))
    trav ρ (!-E (M ✴⟨ sp+ ⟩ N)) =
      let ρₗ ↘, sp+′ ,↙ ρᵣ = env-+ (ρ , sp+) in
      !-E (trav ρₗ M ✴⟨ sp+′ ⟩ trav (bindEnv ρᵣ) N)
\end{code}
%</trav>

%<*dlv>
\begin{code}
  ↙ᵛ : Weakening _∋_
  ↙ᵛ Q0 v .idx = ↙ (v .idx)
  ↙ᵛ Q0 v .tyq = v .tyq
  ↙ᵛ Q0 v .basis = v .basis ++ₙ Q0
\end{code}
%</dlv>

\begin{code}
  open Kit
\end{code}

%<*var-kit>
\begin{code}
  ∋-kit : Kit _∋_
  ∋-kit = kit ↙ᵛ id var

  ren = trav ∋-kit
\end{code}
%</var-kit>

\begin{code}
  instance
    identityEnv-⊢ : IdentityEnv _⊢_
    identityEnv-⊢ .pure = var
\end{code}

%<*dlv-env>
\begin{code}
  ↙^Env : {𝓥 : Ctx → Ty → Set} → Weakening 𝓥 → Weakening [ 𝓥 ]_⇒ᵉ_
  ↙^Env ↙^𝓥 Q0 ρ .M = [ ρ .M │ 0ᴹ ]
  ↙^Env ↙^𝓥 Q0 ρ .asLinRel = [ ρ .asLinRel │ 0AsLinRel ]AsLinRel
  ↙^Env ↙^𝓥 Q0 ρ .sums = ρ .sums , ≤*→0* Q0
  ↙^Env ↙^𝓥 Q0 ρ .lookup (r , sp0) v = ↙^𝓥 (0*→≤* sp0) (ρ .lookup r v)

  ↙ᵛ-env :
    ∀ {s t P} {γ : Vector Ty (s <+> t)} → P ∘ ↘ ≤* 0* →
    [ _∋_ ] ctx P γ ⇒ᵉ ctx (P ∘ ↙) (γ ∘ ↙)
  ↙ᵛ-env Q0 = ↙^Env ↙ᵛ Q0 id^Env
\end{code}
%</dlv-env>

%<*sub>
\begin{code}
  ⊢-kit : Kit _⊢_
  ⊢-kit = kit (λ Q0 → ren (↙ᵛ-env Q0)) var id

  sub = trav ⊢-kit
\end{code}
%</sub>

\begin{code}
  idˢ : ∀ {Γ} → [ _⊢_ ] Γ ⇒ᵉ Γ
  idˢ = id^Env

  psh-⊢ : IsPresheaf _⊢_
  psh-⊢ QP = ren (subuse^Env QP)

  mutual
    open With-psh^𝓥 psh-⊢
\end{code}

%<*subSingle>
\begin{code}
    sub[-] : ∀ {s P Q R γ r A B} → R ≤[ Q +*_] ◇ _≤[ r *ₗ P ] →
      ctx {s} P γ ⊢ A → ctx Q γ ++ᶜ [ r · A ]ᶜ ⊢ B → ctx R γ ⊢ B
    sub[-] {Q = Q} {R} {γ} {r} {A} (sp+ , sp*) t = sub σ
      where
      σ : [ _⊢_ ] ctx R γ ⇒ᵉ ctx Q γ ++ᶜ [ r · A ]ᶜ
      σ = ++ᵉ (idˢ ✴⟨ sp+ ⟩ [-]ᵉ (⟨ sp* ⟩· t))
\end{code}
%</subSingle>