### 2021-09-07 tex/cpp21/quant-framework.tex

\documentclass[sigplan,10pt,anonymous,review]{acmart}
\settopmatter{printfolios=true,printccs=false,printacmref=false}

\usepackage[conor]{agda}
\usepackage{catchfilebetweentags}
\usepackage{cleveref}
\usepackage{ebproof}
\usepackage{mathrsfs}
\usepackage{mathtools}
\usepackage{newunicodechar}
\usepackage{stmaryrd}
\usepackage{todonotes}

\input{../macros.tex}
\input{lagda.tex}

\def\Syntaxtex{../../generic-lr/src/latex/Generic/Linear/Syntax.tex}
\def\Interpretationtex%
{../../generic-lr/src/latex/Generic/Linear/Syntax/Interpretation.tex}

\begin{document}

\title{A Framework for Substructural Programming Languages}

\author{James Wood}
\orcid{0000-0002-8080-3350}
\affiliation{%
  \institution{University of Strathclyde}%
  \city{Glasgow}%
  \country{UK}%
}
\email{james.wood.100@strath.ac.uk}

\author{Robert Atkey}
\orcid{0000-0002-4414-5047}
\affiliation{%
  \institution{University of Strathclyde}%
  \city{Glasgow}%
  \country{UK}%
}
\email{robert.atkey@strath.ac.uk}

\maketitle

\section{Introduction}\label{sec:intro}

\section{Vectors over semirings}\label{sec:algebra}
\input{algebra}

\section{Environments}\label{sec:env}
\input{env}

\section{Generic syntax}\label{sec:syntax}
\input{syntax}

\section{Semantics}\label{sec:semantics}

\section{Conclusions}\label{sec:conc}

\end{document}