### 2021-09-07 tex/cpp21/syntax.tex

\subsection{Descriptions}

We capture the form of rules exemplified previously\todo{Previously?} via
\emph{descriptions} of rules.
The key to making these descriptions work is that they only allow syntactic
forms that preserve environments.
These forms are: absence and multiplicity of subterms with the same usage
annotations, absence and multiplicity of subterms with summed usage annotations,
scaling of a subterm, and variable binding.

A type \AgdaRecord{System} is made up of multiple \AgdaRecord{Rule}s.
Each \AgdaRecord{Rule} comprises a tree of \AgdaDatatype{Premises} and a type
of conclusion.

\ExecuteMetaData[\Syntaxtex]{Scoped}

A single premise is introduced by the
\AgdaInductiveConstructor{$\langle$\_`$\vdash$\_$\rangle$} constructor.
This allows binding of some variables \AgdaBound{$\Gamma$} (with specified
types and usage annotations) and the specification of a conclusion type
\AgdaBound{A}.
The rest of the constructors are descriptions for the first-order bunched
connectives, and are interpreted directly as such.
\todo{Remove Box}
\todo{Remove instance arguments}

\ExecuteMetaData[\Syntaxtex]{Premises}
\ExecuteMetaData[\Interpretationtex]{semp}

\ExecuteMetaData[\Syntaxtex]{Rule}
\ExecuteMetaData[\Interpretationtex]{semr}

\ExecuteMetaData[\Syntaxtex]{System}
\ExecuteMetaData[\Interpretationtex]{sems}