### 2021-09-14 tex/prelim.tex

\chapter{Mathematical preliminaries}

Categories, (co)products, monoidal categories, string diagrams (in Rel),
monoid objects.
Examples: Set, Rel, partial functions, commutative monoids.