### 2021-09-14 tex/simple.tex

\chapter{Mechanisation of simple types}\label{sec:simple}

\section{Term representation}
\input{terms.tex}
\section{Renaming and substitution}\label{sec:kits}
\input{kits.tex}
\section{Generic semantics}
\input{gen-sem.tex}
\section{Generic syntax}
\section{Natural deduction vs sequent calculus}
\input{ndsc.tex}