### 2021-09-14 tex/wrel.tex

\chapter{Semantics in worldly relations}\label{sec:wrel}