### 2021-09-14 tex/cpp21/quant-framework.tex

\documentclass[sigplan,10pt,anonymous,review]{acmart}
\settopmatter{printfolios=true,printccs=false,printacmref=false}

\usepackage[conor]{agda}
\usepackage{catchfilebetweentags}
\usepackage{cleveref}
\usepackage{cmll}
\usepackage{ebproof}
\usepackage{mathrsfs}
\usepackage{mathtools}
\usepackage{newunicodechar}
\usepackage{stmaryrd}
\setlength{\marginparwidth}{2cm}
\usepackage{todonotes}
\usepackage{turnstile}

\input{../macros.tex}
\input{lagda.tex}

\def\genericlr{../../generic-lr/src/processed-latex}
\def\Syntaxtex{\genericlr/Generic/Linear/Syntax.tex}
\def\Interpretationtex{\genericlr/Generic/Linear/Syntax/Interpretation.tex}
\def\Maptex{\genericlr/Generic/Linear/Syntax/Interpretation/Map.tex}
\def\Termtex{\genericlr/Generic/Linear/Syntax/Term.tex}
\def\Semanticstex{\genericlr/Generic/Linear/Semantics.tex}
\def\Renamingtex{\genericlr/Generic/Linear/Renaming.tex}
\def\PaperExamplestex{\genericlr/Generic/Linear/Example/PaperExamples.tex}

\begin{document}

\title{A Framework for Substructural Type Systems}

\author{James Wood}
\orcid{0000-0002-8080-3350}
\affiliation{%
  \institution{University of Strathclyde}%
  \city{Glasgow}%
  \country{UK}%
}
\email{james.wood.100@strath.ac.uk}

\author{Robert Atkey}
\orcid{0000-0002-4414-5047}
\affiliation{%
  \institution{University of Strathclyde}%
  \city{Glasgow}%
  \country{UK}%
}
\email{robert.atkey@strath.ac.uk}

\keywords{substructural, linear, metatheory, type system, mechanisation}

\begin{abstract}
  We present a framework for mechanising the metatheory of certain substructural
  type systems.
\end{abstract}

\maketitle

\section{Introduction}\label{sec:intro}
\input{intro}

\section{Vectors over semirings}\label{sec:algebra}
\input{algebra}

\section{Generic syntax}\label{sec:syntax}
\input{syntax}

\section{Environments}\label{sec:env}
\input{env}

\section{Semantics}\label{sec:semantics}
\input{semantics}

\section{Conclusions}\label{sec:conc}

\bibliographystyle{ACM-Reference-Format}
\citestyle{acmauthoryear}
\bibliography{quant-framework}

\end{document}