### 2021-09-14 tex/cpp21/semantics.tex

Our aim with the following traversals is to derive a recursion scheme for
terms which abstracts over some of the details of variable binding and
changes to usage annotations.
We follow the structure given by Allais et.\ al.~\cite{AACMM20} closely, with
the main difference having come already in the syntax and environments.

\subsection{Renameability}

{\color{red}(Maybe this goes in the Environment section.)}
We saw in \cref{sec:env} that our notion of \emph{environment} specialises to
\emph{simultaneous renaming} by setting $\V$ to be $\sqni$.

\ExecuteMetaData[\Renamingtex]{Ren}
\ExecuteMetaData[\Renamingtex]{Boxr}
\ExecuteMetaData[\Renamingtex]{Renameable}

\subsection{Syntactic traversal}

\subsection{Semantic traversal}

\ExecuteMetaData[\Semanticstex]{Kripke}
\ExecuteMetaData[\Semanticstex]{Semantics}