### 2021-09-17 tex/cpp21/quant-framework.tex

\documentclass[sigplan,10pt,anonymous,review]{acmart}
\settopmatter{printfolios=true,printccs=false,printacmref=false}

\citestyle{acmauthoryear}
\bibliographystyle{ACM-Reference-Format}

\usepackage[conor]{agda}
\usepackage{catchfilebetweentags}
\usepackage{cleveref}
\usepackage{cmll}
\usepackage{ebproof}
\usepackage{mathrsfs}
\usepackage{mathtools}
\usepackage{newunicodechar}
\usepackage{stmaryrd}
\setlength{\marginparwidth}{2cm}
\usepackage{todonotes}
\usepackage{turnstile}

\input{../macros.tex}
\input{lagda.tex}

\def\genericlr{../../generic-lr/src/processed-latex}
\def\Syntaxtex{\genericlr/Generic/Linear/Syntax.tex}
\def\Interpretationtex{\genericlr/Generic/Linear/Syntax/Interpretation.tex}
\def\Maptex{\genericlr/Generic/Linear/Syntax/Interpretation/Map.tex}
\def\Termtex{\genericlr/Generic/Linear/Syntax/Term.tex}
\def\Semanticstex{\genericlr/Generic/Linear/Semantics.tex}
\def\Syntactictex{\genericlr/Generic/Linear/Semantics/Syntactic.tex}
\def\Renamingtex{\genericlr/Generic/Linear/Renaming.tex}
\def\PaperExamplestex{\genericlr/Generic/Linear/Example/PaperExamples.tex}
\def\Snippetstex{../../agda/processed-latex/Snippets.tex}

\begin{document}

\title{A Framework for Substructural Type Systems}

\author{James Wood}
\orcid{0000-0002-8080-3350}
\affiliation{%
  \institution{University of Strathclyde}%
  \city{Glasgow}%
  \country{UK}%
}
\email{james.wood.100@strath.ac.uk}

\author{Robert Atkey}
\orcid{0000-0002-4414-5047}
\affiliation{%
  \institution{University of Strathclyde}%
  \city{Glasgow}%
  \country{UK}%
}
\email{robert.atkey@strath.ac.uk}

\keywords{substructural, linear, metatheory, type system, mechanisation}

\begin{abstract}
  Mechanisation of programming language research is of growing interest, and
  the act of mechanising type systems and their metatheory is generally becoming
  easier as new techniques are invented.
  However, state-of-the-art techniques mostly rely on \emph{structurality} of
  the type system --- that weakening, contraction, and exchange are admissible
  and variables can be used unrestrictedly once assumed.
  Linear logic, and many related subsequent systems, provide motivations for
  breaking some of these assumptions.

  We present a framework for mechanising the metatheory of certain
  substructural type systems, in a style resembling mechanised metatheory of
  structural type systems.
  The framework covers a wide range of simply typed syntaxes with semiring
  usage annotations, via a metasyntax of typing rules.
  The metasyntax for the premises of a typing rule is related to bunched logic,
  featuring both sharing and separating conjunction, roughly corresponding to
  the additive and multiplicative features of linear logic.
  The bunched flavour is carried over into the semantics, together with the use
  of basic linear algebra constructs.
  For example, \emph{environments} are presented equivalently as values
  accumulated via separating conjunction, and as functions from variables to
  values supported by linear maps.
  Producing a generic semantic traversal has us combine environments with a
  separating implication, producing a Kripke function space of the form
  $\Box(A \wand B)$.
  From the generic semantic traversal, we derive totally generic renaming and
  substitution operations, a specific denotational semantics, and a
  syntax-generic \emph{usage elaborator} which greatly facilitates writing
  concrete terms.
\end{abstract}

\maketitle

\section{Introduction}\label{sec:intro}
\input{intro}

\section{Vectors over semirings}\label{sec:algebra}
\input{algebra}

\section{Generic syntax}\label{sec:syntax}
\input{syntax}

\section{Environments}\label{sec:env}
\input{env}

\section{Semantics}\label{sec:semantics}
\input{semantics}

\section{Example traversals}\label{sec:examples}
\input{examples}

\section{Conclusions}\label{sec:conc}

\bibliography{quant-framework}

\end{document}