### 2021-10-05 tex/esop22/semantics.tex
[ writing ] explaining semantics

Our aim with the following traversals is to derive a recursion scheme for
terms which abstracts over some of the details of variable binding and
changes to usage annotations.
We follow the structure given by \citet{AACMM20} closely, with
the main difference having come already in the syntax and environments.

\subsection{Renameability}

{\color{red}(Maybe this goes in the Environment section.)}
We saw in \cref{sec:env} that our notion of \emph{environment} specialises to
\emph{simultaneous renaming} by setting $\V$ to be $\sqni$.

\ExecuteMetaData[\Renamingtex]{Ren}
\ExecuteMetaData[\Renamingtex]{Boxr}
\ExecuteMetaData[\Renamingtex]{Renameable}

\begin{lemma}[\AgdaFunction{ren\textasciicircum{}Env}]\label{thm:env-ren}
  If $\plr{-} \sdtstile{}\V A$ is renameable for all $A$, then so is
  $\plr{-} \env\V \Delta$ for all $\Delta$.
\end{lemma}
\begin{proof}
  For renameable $\V$, we can implement $\mathrm{lift}$ as in
  \cref{thm:env-comp} with $\U \coloneqq {\sqni}$ and
  $\V \coloneqq \W \coloneqq \V$.
  Renaming the environment then becomes an instance of environment composition.
\end{proof}

\subsection{Semantic traversal}

Earlier, when talking about syntax, we saw \AgdaFunction{Scope} as a simple way
to let an open family admit context extensions.
However, a key component of our generic semantic traversal is to make use of
the open family \AgdaBound{$\V$} of \emph{values}, which are the sort of thing
we store in an environment.
The definition \AgdaFunction{Kripke} gives an alternative to
\AgdaFunction{Scope} which interprets the newly bound variables via a
requirement of $\V$-values rather than extra assumptions for the
$\C$-computation.

\ExecuteMetaData[\Semanticstex]{Kripke}

\AgdaFunction{Wrap} is a device that turns any type family into an equivalent
type family that is definitionally injective in its indices.
It turns the type family into a parametrised record with a single field
\AgdaField{get} whose type is the type in the body of the $\lambda$-abstraction.
For understanding the meaning of the type family, \AgdaFunction{Wrap} can be
ignored.
If $\Delta$ is of the form $\gr{s_1}B_1, \ldots, \gr{s_n}B_n$, then
\ExecuteMetaData[\Snippetstex]{KripkeVCDGA}\ is equivalent to
\ExecuteMetaData[\Snippetstex]{KripkeExpanded}.
That is to say, the Kripke function is expecting a value for each newly bound
variable, at the multiplicity of its annotation, together with the resources
supporting each of those values.
The need for \AgdaFunction{$\Box^r$} is exemplified by the following construct
\AgdaFunction{reify}, where we weaken \AgdaBound{$\Gamma$} by a $\gr0$ed-out
version of \AgdaBound{$\Delta$}.
The \AgdaBound{$\Delta$} then gets filled in by the $\V$-values.

For typical choices of \AgdaBound{$\V$} (specifically, those which support
identity environments as per \cref{thm:env-id}),
\AgdaFunction{Kripke}\AgdaSpace{}\AgdaBound{$\V$}\AgdaSpace{}\AgdaBound{$\C$}
is stronger than \AgdaFunction{Scope}\AgdaSpace{}\AgdaBound{$\C$}.
To show this, we instantiate the Kripke function with the renaming
$\swarrow^r : \Gamma, \gr0\delta \env\sqni \Gamma$ to extend the scope, and
pass it an argument environment
$\plr{\id^\V; \searrow^r} : \gr0\gamma, \Delta \env\V \Delta$ to fill in the
extended part.
Post-composition of a renaming comes from \cref{thm:env-postren}.

\ExecuteMetaData[\Syntactictex]{reify}

We tend to use \AgdaFunction{reify} in syntactic traversals, where, for example,
\ExecuteMetaData[\Snippetstex]{term-with-bound}\ could be passed to a syntax
constructor and make sense.
\todo{It's also used in the denotational semantics, so we need a better
  explanation.}

\ExecuteMetaData[\Semanticstex]{Semantics}

%\ExecuteMetaData[\Semanticstex]{Comp}

We mutually define the action \AgdaFunction{semantics} and its lemma
\AgdaFunction{body}.
The purpose of \AgdaFunction{semantics} is to turn a term into a
\AgdaBound{$\C$}-value using a \AgdaBound{$\V$}-environment and the fields of
\AgdaRecord{Semantics}.
Meanwhile, \AgdaFunction{body} does a similar job, but also deals with
newly bound variables.
In particular, \AgdaFunction{body} takes a term in a context extended by
\AgdaBound{$\Theta$}, and produces a Kripke function from
\AgdaBound{$\V$}-values for \AgdaBound{$\Theta$} to \AgdaBound{$\C$}-values.

\ExecuteMetaData[\Semanticstex]{semantics-type}

In \AgdaFunction{semantics}, to deal with a variable \AgdaBound{v}, we look it
up in the environment \AgdaBound{$\rho$}, then use the
\AgdaField{$\sem{\text{var}}$} field to map the resulting
\AgdaBound{$\V$}-value to a \AgdaBound{$\C$}-value.
To deal with a syntactic construct, we use the functoriality of a layer of
syntax, together with recursive calls to \AgdaFunction{semantics} via
\AgdaFunction{body}.
Recall that \AgdaFunction{relocate} (\cref{thm:env-resize}) adjusts the
environment \AgdaBound{$\rho$} to work in the usage contexts of the subterms.

\ExecuteMetaData[\Semanticstex]{semantics}

For \AgdaFunction{body}, we are given
$\rho : \grP\gamma \env\V \Delta$,
$\mathit{th} : \grP\gr{^+}\gamma^+ \env\sqni \grP\gamma$,
$\sigma : \grQ\gr{^+}\gamma^+ \env\V \Theta$,
with $r : \grR\gr{^+} \leq \grP\gr{^+} + \grQ\gr{^+}$.
\todo{This needs some explanation.}
We are aiming for an environment of type
$\grR\gr{^+}\gamma \env\V \Delta, \Theta$, which we get by renaming $\rho$ by
$\mathit{th}$ to give $\rho'$, then using \cref{thm:env-rules} to stick
together the environment $\rho'$ for old variables $\Delta$ and the
environment $\sigma$ for the new variables $\Theta$.

\ExecuteMetaData[\Semanticstex]{body}