### 2021-10-14 tex/esop22/quant-framework.tex

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}

\usepackage[conor]{agda}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{catchfilebetweentags}
\usepackage{cleveref}
\usepackage{cmll}
\usepackage{ebproof}
\usepackage{mathrsfs}
\usepackage{mathtools}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{newunicodechar}
\usepackage{stmaryrd}
\usepackage{todonotes}
\usepackage{turnstile}
\usepackage{mathpartir}

\input{../macros.tex}
\input{lagda.tex}

\def\genericlr{../../generic-lr/src/processed-latex}
\def\Syntaxtex{\genericlr/Generic/Linear/Syntax.tex}
\def\Interpretationtex{\genericlr/Generic/Linear/Syntax/Interpretation.tex}
\def\Maptex{\genericlr/Generic/Linear/Syntax/Interpretation/Map.tex}
\def\Termtex{\genericlr/Generic/Linear/Syntax/Term.tex}
\def\Semanticstex{\genericlr/Generic/Linear/Semantics.tex}
\def\Syntactictex{\genericlr/Generic/Linear/Semantics/Syntactic.tex}
\def\Renamingtex{\genericlr/Generic/Linear/Renaming.tex}
\def\PaperExamplestex{\genericlr/Generic/Linear/Example/PaperExamples.tex}
\def\HeavyItex{\genericlr/Generic/Linear/Example/PaperExamples/HeavyI.tex}
\def\UsageChecktex{\genericlr/Generic/Linear/Example/UsageCheck.tex}
\def\WReltex{\genericlr/Generic/Linear/Example/WRel.tex}
\def\Monotonicitytex{\genericlr/Generic/Linear/Example/WRel/Monotonicity.tex}
\def\MuMuTildetex{\genericlr/Generic/Linear/Example/MuMuTilde.tex}
\def\MuMuTildeTermtex{\genericlr/Generic/Linear/Example/MuMuTilde/Term.tex}
\def\Snippetstex{../../agda/processed-latex/Snippets.tex}

\newcommand{\figref}[1]{Figure \ref{#1}}

\begin{document}
%
\title{A Framework for Substructural Type Systems
  \thanks{Supported by organization x.}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Anonymous authors}
%\author{James Wood\inst{1}\orcidID{0000-0002-8080-3350} \and
%  Robert Atkey\inst{1}\orcidID{0000-0002-4414-5047}}
%
\authorrunning{Anon.}
%\authorrunning{J. Wood, R. Atkey}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
%\institute{University of Strathclyde, Glasgow, UK
%\email{\{james.wood.100,robert.atkey\}@strath.ac.uk}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
  Mechanisation of programming language research is of growing interest, and
  the act of mechanising type systems and their metatheory is generally becoming
  easier as new techniques are invented.
  However, state-of-the-art techniques mostly rely on \emph{structurality} of
  the type system --- that weakening, contraction, and exchange are admissible
  and variables can be used unrestrictedly once assumed.
  Linear logic, and many related subsequent systems, provide motivations for
  breaking some of these assumptions.

  We present a framework for mechanising the metatheory of certain
  substructural type systems, in a style resembling mechanised metatheory of
  structural type systems.
  The framework covers a wide range of simply typed syntaxes with semiring
  usage annotations, via a metasyntax of typing rules.
  The metasyntax for the premises of a typing rule is related to bunched logic,
  featuring both sharing and separating conjunction, roughly corresponding to
  the additive and multiplicative features of linear logic.
  We use the uniformity of syntaxes to derive type system-generic renaming,
  substitution, and a form of linearity checking.
  %The bunched flavour is carried over into the semantics, together with the use
  %of basic linear algebra constructs.
  %For example, \emph{environments} are presented equivalently as values
  %accumulated via separating conjunction, and as functions from variables to
  %values supported by linear maps.
  %Producing a generic semantic traversal has us combine environments with a
  %separating implication, producing a Kripke function space of the form
  %$\Box(A \wand B)$.
  %From the generic semantic traversal, we derive totally generic renaming and
  %substitution operations, a specific denotational semantics, and a
  %syntax-generic \emph{usage elaborator} which greatly facilitates writing
  %concrete terms.

\keywords{First keyword  \and Second keyword \and Another keyword.}
\end{abstract}
%
%
%
\begin{enumerate}
\item (Section 1)
  \begin{enumerate}
  \item   Motivate intrinsically typed syntax (in Agda)
    \begin{enumerate}
    \item Show the ``GADT'' for well scoped syntax
    \item Show the well typed version
    \end{enumerate}
  \item Motivate adding posemiring annotations
    \begin{enumerate}
    \item Pointing to the existing work (Brunel et al., Ghica and Smith, Granule, previous of ours, Bernardy and Abel)
    \item The fact that we can encode lots of other systems (DILL, PD, monotonicity, information flow, differential privacy)
    \end{enumerate}
  \end{enumerate}
\item (Section 2) Background / notation fixing for:
  \begin{enumerate}
  \item Partially ordered semirings
  \item Vectors
  \item Linear maps and linear relations (as a generalisation of $P \leq Q\Psi$ which is easier to work with)
  \end{enumerate}
\item (Section 3) Bunched Typing Rules
  \begin{enumerate}
  \item Introduces $\lambda\mathcal{R}$
  \item Rephrases the typing rules in terms of bunches (as well as implicit contexts)
  \end{enumerate}
\item (Section 4) Universe of Generic Usage-annotated syntax
  \begin{enumerate}
  \item Premises
  \item Rules
  \item Systems
  \item Example(s)
  \item Intepretation as Terms
  \item Examples of Terms; using pattern synonyms (point out that proofs of usage accounting are painful)
  \end{enumerate}
\item (Section 5) Example Syntaxes
  \begin{enumerate}
  \item $\lambda\mathcal{R}$ (incl DILL, PD, differential privacy, monotonicity)
  \item Inductive types with a well resourced fold
  \item Bernardy-Abel style
  \item First-order (circuitry)
  \item LNL
  \item $\lambda\mu\tilde{\mu}$
  \end{enumerate}
\item (Section 6) Environments
  \begin{enumerate}
  \item Motivate in terms of renaming and substitution for simple types (where it is just a function)
  \item Motivate the general idea of using linear relations (i.e. transformations + weakening)
    \begin{enumerate}
    \item Key idea / slogan : environment = change of basis
    \end{enumerate}
  \item Definition of environments
  \item Examples
  \item Properties, incl. preservation of linear operations
  \end{enumerate}
\item (Section 7) Semantics, or How to apply an Environment to a Term
  \begin{enumerate}
  \item Renameability
  \item $\mathsf{semantics}$
  \end{enumerate}
\item (Section 8) Examples of Semantics
  \begin{enumerate}
  \item Renaming and Substitution (term-valued)
  \item Denotational Semantics (relational world-indexed (\texttt{Generic/Linear/Example/WRel.agda}))
  \item Usage Elaborator
  \end{enumerate}
\item (Section 9) Related Work
\item (Section 10) Conclusion
\end{enumerate}

\section{Introduction}\label{sec:intro}
\input{intro}

\section{Vectors over semirings}\label{sec:algebra}
\input{algebra}

\section{Bunched Typing Rules}\label{sec:bunched-rules}
\input{bunchedrules}

\section{Generic syntax}\label{sec:syntax}
\input{syntax}

\section{Environments}\label{sec:env}
\input{env}

\section{Semantics}\label{sec:semantics}
\input{semantics}

\section{Example traversals}\label{sec:examples}
\input{examples}

\section{Conclusions}\label{sec:conc}
\input{conc}
%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{splncs04nat}
\bibliography{quant-framework}
\end{document}