### 2021-10-26 tex/framework.tex
[ dump ] framework work from the paper

\chapter{A framework for usage-restricted calculi}

\section{Syntax}

\input{syntax}

\section{Semantics}

\input{semantics}