### 2022-05-27 tex/prelim.tex

\chapter{Mathematical preliminaries}

Categories, (co)products, monoidal categories, string diagrams (in Rel),
monoid objects.
Examples: Set, Rel, partial functions, commutative monoids.

\begin{definition}
  A \emph{category} is a collection $\obj$ of objects, a family
  $\hom : \obj \times \obj \to \Set$, for each object $A$ an element
  $\id : \hom(A,A)$, and for each triple of objects $A, B, C$ a function
  ${\circ} : \hom(B,C) \times \hom(A,B) \to \hom(A,C)$ such that the following
  laws hold for any $f : \hom(A,B), g : \hom(B,C), h : \hom(C,D)$.
  \begin{itemize}
    \item $\id \circ f = f$
    \item $f \circ \id = f$
    \item $(h \circ g) \circ f = h \circ (g \circ f)$
  \end{itemize}
  I sometimes write $f; g$ for $g \circ f$.
\end{definition}

I give a somewhat unusual definition of \emph{monoidal categories} via
\emph{multicategories}.
I use this approach to capture the purpose of monoidal categories in many
examples, particularly in algebra and programming languages: to provide a way to
encode morphisms of arbitrary arity as unary morphisms by representing $n$-ary
domains as internal $n$-ary products.

I choose to present multicategories in a style where composition models
simultaneous substitution rather than single substitution.
This choice is consistent with the rest of this thesis, which concerns
simultaneous substitution.

\begin{definition}
  A \emph{multicategory} is a collection $\obj$ of objects, a family
  $\hom : \List \obj \times \obj \to \Set$ (with $\hom([A_1, \ldots, A_n], B)$
  written $\hom(A_1, \ldots, A_n; B)$), for each object $A$ an element
  $\id : \hom(A; A)$, and for each list of lists $\Psi$, each list $\Gamma$, and
  each object $A$, a function
  ${\circ} : \hom(\Gamma; A) \times \plr{\prod_i \hom(\Psi_i; \Gamma_i)} \to
  \hom\plr{\sum_i \Psi_i; A}$ such that the following laws hold.
  \begin{itemize}
    \item $\id \circ f = f$
    \item $f \circ \plr{\lambda i.~\id} = f$
    \item $h \circ \plr{\lambda i.~g_i \circ \lambda j.~f_{ij}} =
      (h \circ \lambda i.~g_i) \circ \lambda ij.~f_{ij}$
  \end{itemize}
\end{definition}

\begin{example}
  The following give rise to multicategories:
  \begin{itemize}
    \item Sets and multi-ary functions (functions of the form
      $A_1 \times \cdots \times A_n \to B$)
    \item $K$-vector spaces and multilinear maps
    \item For any simple type theory with sequents of the form
      $\Gamma \vdash A$, where $\Gamma$ is a list of types and $A$ is a type,
      types and open terms (up to $\beta\eta$-equivalence) form a multicategory.
  \end{itemize}
\end{example}

Related to the multicategory of $K$-vector spaces and multilinear maps is the
multicategory of commutative monoids and multilinear maps.
I use the latter later in this chapter\todo{forward reference}, so I will
develop it fully as an example.

\begin{example}
  The multicategory of \emph{commutative monoids and multilinear maps} has as
  objects commutative monoids.
  A morphism of type $A_1, \ldots, A_n \to B$ is a function
  $f : A_1 \times \cdots \times A_n \to B$ such that the following laws hold:
  \begin{itemize}
    \item $f(x_1, \ldots, 0_{A_i}, \ldots, x_n) = 0_B$
    \item $f(x_1, \ldots, y +_{A_i} z, \ldots, x_n) =
      f(x_1, \ldots, y, \ldots, x_n) +_B f(x_1, \ldots, z, \ldots, x_n)$
  \end{itemize}
  The identity function of type $A \to A$ serves as the identity
  (multi)morphism.
  Composition is given by multi-place composition of functions, which satisfies
  the above laws because in a composite $g \circ [f_1, \ldots, f_n]$, a
  commutative monoid operation in the domain of the composite bubbles through
  one of the $f_i$ to become one operation in position $i$ of the domain of $g$,
  and then bubbles through $g$.
  That this identity and composition follow the equational laws follows
  straightforwardly by considering the underlying functions.
\end{example}

\begin{definition}
  A multicategory $\C$ gives rise to its \emph{category of contexts} $\C^*$ as
  follows:
  \begin{itemize}
    \item $\obj(\C^*) \coloneqq \List(\obj\C)$
    \item $\C^*\plr{\sum_i \Psi_i, \Delta} \coloneqq \prod_i \C(\Psi_i; \Delta_i)$
      for any $\Psi : \List(\List(\obj\C))$
    \item $\id^* : \Gamma \to^* \Gamma \coloneqq \overline\id$
    \item $(\sigma : \Delta \to^* \Theta) \circ^* (\rho : \Gamma \to^* \Delta) :
      (\Gamma \to^* \Theta) \coloneqq \lambda i.~\sigma_i \circ \lambda j.~\rho_{ij}$
      \todo{Still wrong}
  \end{itemize}
\end{definition}

\begin{definition}
  Each multicategory $\C$ \emph{restricts} to a category $\C^1$, where
  $\C^1(A, B) \coloneqq \C(A; B)$ and identity and composition are as in $\C$.
\end{definition}

\begin{definition}\label{def:tensor-product}
  A \emph{tensor product} in a multicategory $\C$ of objects $A$ and $B$ is an
  object $A \otimes B$ and a map ${\otimes} : A, B \to A \otimes B$ such that
  any morphism $f : \Gamma, A, B, \Delta \to C$ uniquely factors as:

  \begin{tikzcd}
    \Gamma, A, B, \Delta \arrow[d,"f"']
    \arrow[r,"{\overline\id, \otimes, \overline\id}"]
    & \Gamma, A \otimes B, \Delta
    \arrow[dl,dashed,"f^{\Gamma, A \otimes B, \Delta}"] \\
    C
  \end{tikzcd}

  When it is clear from context which two objects are being multiplied, I will
  write just $f^{\otimes}$ for the unique morphism.
\end{definition}

\begin{definition}\label{def:tensor-unit}
  A \emph{tensor unit} in a multicategory $\C$ is an object $I$ and a map
  $I : {\cdot} \to I$ such that any morphism
  $f : \Gamma, \Delta \to C$ uniquely factors as:

  \begin{tikzcd}
    \Gamma, \Delta \arrow[d,"f"']
    \arrow[r,"{\overline\id, I, \overline\id}"]
    & \Gamma, I, \Delta \arrow[dl,dashed,"f^{\Gamma, I, \Delta}"] \\
    C
  \end{tikzcd}

  When it is clear from context where the $I$ is being introduced, I will
  write just $f^I$ for the unique morphism.
\end{definition}

\begin{lemma}\label{thm:tensor-product-terms}
  For each morphism $f : \Gamma \to A$ and $g : \Delta \to B$, we have a
  morphism $(f, g) : \Gamma, \Delta \to A \otimes B$ such that we have, for any
  $h : \Gamma', A, B, \Delta' \to C$, that
  $h^{\otimes} \circ [\overline\id, (f, g), \overline\id] =
  h \circ [\overline\id, f, g, \overline\id]$, and that
  $\id = (\id, \id)^{\otimes} : A \otimes B \to A \otimes B$.
\end{lemma}
\begin{proof}
  The composite $(\otimes) \circ [f, g]$ has the required type.

  To prove the first equation, note that
  \begin{align*}
    h^{\otimes} \circ [\overline\id, (f, g), \overline\id]
    &= h^{\otimes} \circ [\overline\id, {\otimes} \circ [f, g], \overline\id] \\
    &= h^{\otimes} \circ [\overline\id, {\otimes}, \overline\id]
    \circ [\overline\id, f, g, \overline\id].
  \end{align*}
  Then, we have that
  $h^{\otimes} \circ [\overline\id, {\otimes}, \overline\id] = h$
  by the defining diagram of the tensor product (\cref{def:tensor-product}).

  To prove the second equation, note that the diagram below commutes.
  Because this is an instance of the diagram in \cref{def:tensor-product},
  uniqueness gives us the desired equation.

  \begin{tikzcd}
    A, B \arrow[d,"{\otimes}"'] \arrow[r,"{\otimes}"]
    & A \otimes B \arrow[dl,"\id"] \\
    A \otimes B
  \end{tikzcd}
\end{proof}

\begin{lemma}\label{thm:tensor-unit-terms}
  For any $h : \Gamma', \Delta' \to C$, we have that
  $h^I \circ [\overline\id, I, \overline\id] =
  h \circ [\overline\id, \overline\id]$
  and $\id = I^I : I \to I$.
\end{lemma}
\begin{proof}
  The first equation is just the commuting diagram of \cref{def:tensor-unit},
  modulo composition with identity morphisms.
  For the second equation, note that the diagram below commutes.
  Because this is an instance of the diagram in \cref{def:tensor-unit},
  uniqueness gives us the desired equation.

  \begin{tikzcd}
    {\cdot} \arrow[d,"I"'] \arrow[r,"I"]
    & I \arrow[dl,"\id"] \\
    I
  \end{tikzcd}
\end{proof}

\begin{lemma}
  In a multicategory $\C$ with all tensor products, $\otimes$ induces a
  bifunctor ${\otimes} : \C, \C \to \C$.
\end{lemma}
\begin{proof}
  The action on objects is given by $\otimes$.
  The action on morphisms $f : A \to A'$ and $g : B \to B'$ is given by
  $(f, g)^{\otimes}$.
  Preservation of identity is already proved in \cref{thm:tensor-product-terms},
  and preservation of composition is left as an exercise.
\end{proof}

\begin{lemma}
  For any tensor product and tensor unit, we have the following natural
  isomorphisms:
  \begin{itemize}
    \item $I \otimes A \cong A$
    \item $A \otimes I \cong A$
    \item $(A \otimes B) \otimes C \cong A \otimes (B \otimes C)$
  \end{itemize}
\end{lemma}

\begin{definition}\label{def:monoidal-category}
  A \emph{monoidal category} is a multicategory equipped with a tensor product
  and tensor unit.
  %A \emph{monoidal category} is a category $\C = (\obj, \hom, \id, \circ)$ with
  %functors $I : 1 \to \C$ and ${\otimes} : \C \times \C \to \C$ together with
  %the following natural isomorphisms:
  %\begin{itemize}
  %  \item $\lambda : (I \otimes {-}) \simeq {-}$
  %  \item $\rho : ({-} \otimes I) \simeq {-}$
  %  \item $\alpha :
  %    (({-} \otimes {-}) \otimes {-}) \simeq ({-} \otimes ({-} \otimes {-}))$
  %\end{itemize}
  %such that the following coherence conditions hold:
  %\begin{itemize}
  %  \item
  %    \begin{tikzcd}
  %      (A \otimes I) \otimes B \arrow[dr,"\rho_A \otimes B"'] \arrow[rr,"\alpha_{A,I,B}"]
  %      && A \otimes (I \otimes B) \arrow[dl,"A \otimes \lambda_B"] \\
  %      & A \otimes B
  %    \end{tikzcd}
  %  \item
  %    \begin{tikzcd}
  %      & (A \otimes B) \otimes (C \otimes D)
  %      \arrow[dr,"\alpha_{A,B,C \otimes D}"] &
  %      \\
  %      ((A \otimes B) \otimes C) \otimes D
  %      \arrow[ur,"\alpha_{A \otimes B,C,D}"]
  %      \arrow[d,"\alpha_{A,B,C} \otimes D"]
  %      &&
  %      A \otimes (B \otimes (C \otimes D))
  %      \\
  %      (A \otimes (B \otimes C)) \otimes D
  %      \arrow[rr,"\alpha_{A,B \otimes C,D}"]
  %      &&
  %      A \otimes ((B \otimes C) \otimes D)
  %      \arrow[u,"A \otimes \alpha_{B,C,D}"]
  %    \end{tikzcd}
  %\end{itemize}
\end{definition}

\begin{remark}
  We can talk about monoidal products on a category $\C$ by finding a
  multicategory which restricts to $\C$ and constructing tensor products in it.
\end{remark}

\begin{definition}
  A \emph{monoid} in a multicategory is an object $M$ and multimorphisms
  $\eta : {\cdot} \to M$ and $\mu : M, M \to M$ such that the following
  equations hold:
  \begin{itemize}
    \item $[\eta, \id]; \mu = \id : M \to M$
    \item $[\id, \eta]; \mu = \id : M \to M$
    \item $[\mu, \id]; \mu = [\id, \mu]; \mu : M, M, M \to M$
  \end{itemize}
\end{definition}

\begin{definition}
  TODO: \emph{symmetric monoidal category}
\end{definition}

\begin{definition}
  TODO: \emph{commutative monoid in a symmetric monoidal category}
\end{definition}

Algebra

\begin{definition}
  A \emph{semiring} is a monoid in the multicategory of commutative monoids and
  multilinear maps.
  Unpacked, this means that we have a set $\mathscr R$ together with elements
  $0$ and $1$ and binary operators $+$ and $\cdot$ (with $\cdot$ usually written
  as juxtaposition) such that the following hold for all
  $x, y, z \in \mathscr R$.
  \begin{itemize}
    \item $0 + x = x$; $x + 0 = x$; $(x + y) + z = x + (y + z)$; $x + y = y + x$
    \item $1x = x$; $x1 = x$; $(xy)z = x(yz)$
    \item $0x = 0$; $x0 = 0$; $(x + y)z = xz + yz$; $x(y + z) = xy + xz$
  \end{itemize}
\end{definition}