### 2022-06-06 tex/framework.tex
[ writing ] introduction to MALL

\chapter{A framework for usage-restricted calculi}\label{sec:framework}

\section{Syntax}

\input{syntax}

\section{Semantics}

\input{semantics}

\section{Example semantics}

\input{examples}