### 2022-06-06 tex/ill.tex

Linear logic is a good starting point for our investigations into usage
restrictions because it is well known, well understood, and exhibits most of the
difficulties we will come across.
Linear logic's lack of general weakening and contraction will be a typical
feature of the calculi we will see, and the exponential modality $\oc$ (bang)
will be the prototype for a range of modalities I will consider.

\subsection{Motivation of linearity}

While classical linear logic is perhaps better understood, for the sake of this
thesis I focus on intuitionistic linear logic.
The syntactic difference between the two is analogous to the difference between
classical and intuitionistic logic: The latter restricts the former by only
allowing one formula on the right of a sequent.
Classical linear logic is not irrelevant to this thesis, and appears as an
instance of the framework presented in \cref{sec:framework}.
However, the connection of this work to intuitionistic linear logic is more
direct, and reflects the basis of existing semiring-annotated calculi on
(intuitionistic) typed $\lambda$-calculi.

Intuitionistic linear logic is often motivated by its sensitivity to resources.
The fact that, in intuitionistic (and classical) logic, \emph{any} hypothesis
can be discarded or duplicated means that propositions cannot be used to model
things that may be ``possessed'' or ``spent'', despite hypotheses often being
described as things we ``have''.
In contrast, linear logic allows propositions to behave like physical objects,
being moved around but not being created or destroyed (a priori).

Programming applications of linearity include stateful protocols
\citep{Wadler12}, mutually exclusive capabilities (TODO: cite),
and mutation (TODO: cite).
In each of these cases, the act of using up a hypothesis allows us to divide
time into ``before'' and ``after'', with lack of duplication avoiding confusion
over which half we are in, and lack of deletion allowing us to know whether the
corresponding action actually happened.
For example, when modelling mutation, a variable refers to a single state of a
mutable value.
The \texttt{write} operation uses up such a variable, making the old state
inaccessible, and returns a new linear value to represent the new state of the
mutable value.
Such a protocol naturally also supports a \emph{freezing} operation, by which we
relinquish the ability to mutate the value in return for an immutable reference
to the value produced.

%In LJ, two distinct, but equiderivable, canonical definitions of $\wedge$ can
%be given.
%
%In the first approach, $\wedge^-$ is conceived to be a
%\emph{categorial product}.
%Such a product is constructed by providing two maps in from the same antecedents
%$\Gamma$.
%It is used by projecting out one of the sides.
%
%\begin{mathpar}
%  \inferrule*[right=$\wedge^-$-r]
%  {\Gamma \vdash A \\ \Gamma \vdash B}
%  {\Gamma \vdash A \wedge^- B}
%
%  \and
%
%  \inferrule*[right=$\wedge^-$-l$_0$]
%  {\Gamma, A \vdash C}
%  {\Gamma, A \wedge^- B \vdash C}
%
%  \and
%
%  \inferrule*[right=$\wedge^-$-l$_1$]
%  {\Gamma, B \vdash C}
%  {\Gamma, A \wedge^- B \vdash C}
%\end{mathpar}
%
%In the second approach, $\wedge^+$ is conceived to be a \emph{tensor product}.
%Intuitively, $\wedge^+$ internalises the comma of context concatenation.
%Such a product is used by giving access to both sides simultaneously.
%It is constructed by constructing both sides and combining the antecedents
%required by each side.
%
%\begin{mathpar}
%  \inferrule*[right=$\wedge^+$-r]
%  {\Gamma \vdash A \\ \Delta \vdash B}
%  {\Gamma, \Delta \vdash A \wedge^+ B}
%
%  \and
%
%  \inferrule*[right=$\wedge^+$-l]
%  {\Gamma, A, B \vdash C}
%  {\Gamma, A \wedge^+ B \vdash C}
%\end{mathpar}
%
%When we prove the logical equivalence of these two formulations, we notice that
%the structural rules of weakening and contraction are essential.
%When we remove weakening and contraction, $\wedge^-$ and $\wedge^+$ become
%logically distinct connectives, which we notate $\&$ and $\otimes$,
%respectively.
%
%\begin{mathpar}
%  \inferrule*[right=$\wedge^-$-r]
%  {%
%    \inferrule*[right=Weak$^*$,fraction={\cdot\cdots\cdot}]
%    {\Gamma \vdash A}
%    {\Gamma, \Delta \vdash A}
%    \\
%    \inferrule*[Right=Weak$^*$,fraction={\cdot\cdots\cdot}]
%    {\Delta \vdash B}
%    {\Gamma, \Delta \vdash B}
%  }
%  {\Gamma, \Delta \vdash A \wedge^- B}
%
%  \and
%
%  \inferrule*[right=Contr]
%  {%
%    \inferrule*[Right=$\wedge^-$-l$_0$]
%    {%
%      \inferrule*[Right=$\wedge^-$-l$_1$]
%      {\Gamma, A, B \vdash C}
%      {\Gamma, A, A \wedge^- B \vdash C}
%    }
%    {\Gamma, A \wedge^- B, A \wedge^- B \vdash C}
%  }
%  {\Gamma, A \wedge^- B \vdash C}
%\end{mathpar}
%
%\begin{mathpar}
%  \inferrule*[right=Contr$^*$,fraction={\cdot\cdots\cdot}]
%  {%
%    \mprset{defaultfraction}
%    \inferrule*[Right=$\wedge^+$-r]
%    {\Gamma \vdash A \\ \Gamma \vdash B}
%    {\Gamma, \Gamma \vdash A \wedge^+ B}
%  }
%  {\Gamma \vdash A \wedge^+ B}
%
%  \and
%
%  \inferrule*[right=$\wedge^+$-l]
%  {%
%    \inferrule*[Right=Weak]
%    {\Gamma, A \vdash C}
%    {\Gamma, A, B \vdash C}
%  }
%  {\Gamma, A \wedge^+ B \vdash C}
%
%  \and
%
%  \inferrule*[right=$\wedge^+$-l]
%  {%
%    \inferrule*[Right=Weak]
%    {\Gamma, B \vdash C}
%    {\Gamma, A, B \vdash C}
%  }
%  {\Gamma, A \wedge^+ B \vdash C}
%\end{mathpar}

\subsection{The multiplicative-additive fragment}

The multiplicative-additive fragment of linear logic (MALL) is the fragment
where all hypotheses are linear (must be used exactly once).
We will extend MALL with the \emph{exponential} modality in
\cref{sec:bang-modality}.
MALL is unable to embed intuitionistic or classical logic (collectively
known as \emph{structural logics}), as MALL is unable to reflect any of the
discarding or duplication that can be done in proofs in structural logics.

In short, the syntax of intuitionistic MALL can be described as intuitionistic
logic with the structural rules of \emph{weakening} and \emph{contraction}
removed.
However, without the presence of weakening and contraction, we have to be more
careful about the rules we state, so as not to accidentally introduce weakening
and contraction admissibly.
The lack of these structural rules also allows us to observe a new phenomenon:
the distinction (at the level of provability) between \emph{additive} and
\emph{multiplicative} formulations of existing connectives (in particular, in
the intuitionistic case, the conjunction connective).

I present MALL in \cref{fig:mall} in a sequent calculus style, as it was
presented by \citet{girard87linear}.

To encode what it means to use a hypothesis \emph{exactly once}, we first need
to decide what counts as a use.
The simplest case is that the identity sequent counts as a single use of its
sole hypothesis, and conversely does not count as a use of any other hypotheses.
For sequential proofs, created by the \TirName{Cut} rule, if we have a proof of
$A$ using $\Gamma$, and a proof of $B$ using $\Delta$ and $A$, then we have a
proof of $B$ transitively using $\Delta$ and $\Gamma$.
The exchange rule Exch says that use is invariant under permutation.

For the logical connectives, we have genuine choices as to what it means to use
them.
Two cases --- disjunction ($\oplus$) and (linear) implication ($\multimap$) ---
are somewhat intuitive from intuitionistic logic.
A canonical proof of a disjunction is a tag and a proof of one of the two
disjuncts.
This suggests that a proof of a disjunction only uses the same hypotheses as
the proof of the disjunct we actually choose, with the other disjunct being
irrelevant.
Correspondingly, when we use a disjunction hypothesis, we will only actually use
one of the cases, so each branch should use the same hypotheses.
For implication, use is sequential like with the \TirName{Cut} rule, and its
left rule is more or less the only choice that allows use of the surrounding
hypotheses.

For conjunction, there are two choices: Either the conjuncts \emph{together} use
all of the hypotheses, or each of the conjuncts \emph{individually} uses all of
the hypotheses.
The former choice gives us the tensor-product ($\otimes$), and the latter choice
gives us the with-product ($\with$).
These products are equivalent up to provability in structural logics but
distinct in linear logic.

\begin{figure}
  \begin{align*}
    A, B, C &\Coloneqq X \mid I \mid A \otimes B \mid A \multimap B
              \mid 0 \mid A \oplus B \mid \top \mid A \with B \\
    \Gamma, \Delta, \Theta &\Coloneqq {\cdot} \mid \Gamma, A
  \end{align*}
  \begin{mathpar}
    \ebrule{%
      \infer0[Id]{A \vdash A}
    }

    \and

    \ebrule{%
      \hypo{\Gamma \vdash A}
      \hypo{\Delta, A \vdash B}
      \infer2[Cut]{\Gamma, \Delta \vdash B}
    }

    \and

    \ebrule{%
      \hypo{\Gamma, B, A, \Delta \vdash C}
      \infer1[Exch]{\Gamma, A, B, \Delta \vdash C}
    }

    \and

    \ebrule{%
      \hypo{\Gamma \vdash C}
      \infer1[$I$-L]{\Gamma, I \vdash C}
    }

    \and

    \ebrule{%
      \infer0[$I$-R]{{\cdot} \vdash I}
    }

    \and

    \ebrule{%
      \hypo{\Gamma, A, B \vdash C}
      \infer1[$\otimes$-L]{\Gamma, A \otimes B \vdash C}
    }

    \and

    \ebrule{%
      \hypo{\Gamma \vdash A}
      \hypo{\Delta \vdash B}
      \infer2[$\otimes$-R]{\Gamma, \Delta \vdash A \otimes B}
    }

    \and

    \ebrule{%
      \hypo{\Gamma \vdash A}
      \hypo{\Delta, B \vdash C}
      \infer2[$\multimap$-L]{\Gamma, \Delta, A \multimap B \vdash C}
    }

    \and

    \ebrule{%
      \hypo{\Gamma, A \vdash B}
      \infer1[$\multimap$-R]{\Gamma \vdash A \multimap B}
    }

    \and

    \ebrule{%
      \infer0[$0$-L]{\Gamma, 0 \vdash C}
    }

    \and

    \text{(no $0$-R)}

    \and

    \ebrule{%
      \hypo{\Gamma, A \vdash C}
      \hypo{\Gamma, B \vdash C}
      \infer2[$\oplus$-L]{\Gamma, A \oplus B \vdash C}
    }

    \and

    \ebrule{%
      \hypo{\Gamma \vdash A_i}
      \infer1[$\oplus$-R$_i$]{\Gamma \vdash A_0 \oplus A_1}
    }

    \and

    \text{(no $\top$-L)}

    \and

    \ebrule{%
      \infer0[$\top$-R]{\Gamma \vdash \top}
    }

    \and

    \ebrule{%
      \hypo{\Gamma, A_i \vdash C}
      \infer1[$\with$-L$_i$]{\Gamma, A_0 \with A_1 \vdash C}
    }

    \and

    \ebrule{%
      \hypo{\Gamma \vdash A}
      \hypo{\Gamma \vdash B}
      \infer2[$\with$-R]{\Gamma \vdash A \with B}
    }
  \end{mathpar}
  \caption{Multiplicative-additive fragment of linear logic}
  \label{fig:mall}
\end{figure}

Implication ($\multimap$) and the tensor-product ($\otimes$, $I$) comprise the
\emph{multiplicative} fragment, while disjunction ($\oplus$, $0$) and the
with-product ($\with$, $\top$) comprise the \emph{additive} fragment.
Categorically, the additive fragment corresponds to products and coproducts,
while the multiplicative fragment corresponds to multicategorical tensor
products and closure.

\subsection{The $\oc$-modality}\label{sec:bang-modality}

\begin{mathpar}
  \ebrule{%
    \hypo{\oc\Gamma \vdash A}
    \infer1[Promotion]{\oc\Gamma \vdash \oc A}
  }

  \and

  \ebrule{%
    \hypo{\Gamma, A \vdash B}
    \infer1[Dereliction]{\Gamma, \oc A \vdash B}
  }

  \and

  \ebrule{%
    \hypo{\Gamma \vdash B}
    \infer1[Weakening]{\Gamma, \oc A \vdash B}
  }

  \and

  \ebrule{%
    \hypo{\Gamma, \oc A, \oc A \vdash B}
    \infer1[Contraction]{\Gamma, \oc A \vdash B}
  }
\end{mathpar}

In the intuitionistic linear logic sequent calculus ILL, $\oc A$ is defined
to be a proposition whose occurrences as antecedents can be deleted
(\TirName{Weakening}) and duplicated (\TirName{Contraction}), from which we can
extract $A$ (\TirName{Dereliction}), and which we can form from a conclusion
$A$ only when all antecedents are of the form $\oc X$ for some proposition $X$
(\TirName{Promotion}).
In short, $\oc A$ can be seen as an intuitionistic version of $A$, supporting
all of the structural rules of LJ, and only being able to be formed when it
does not depend on anything linear.

While this definition of $\oc$ works, in the sense that it gives the intended
class of models and cut elimination is maintained, it has some disadvantages.
Firstly, while the multiplicative and additive connectives are all characterised
by universal properties, $\oc$ is not.
This can be seen by the fact that taking the rules for $\oc$ and replacing each
occurrence of $\oc$ by a fresh connective $\oc'$ produces a logically distinct
connective.
One cannot produce any derivation of $\oc' A \vdash \oc A$ because
\TirName{Promotion} does not apply when there are antecedents not of the form
$\oc X$.
Finally, while $\oc$ is supposed to be a positive connective, it sometimes
behaves like a negative connective.
For example, for a positive connective like $\otimes$, the normal form proof
of the identity sequent $P \otimes Q \vdash P \otimes Q$ starts (from the
bottom) with a left rule and then, with the left in a more amenable form,
applies the right rule.
In contrast, the normal form proof of $\oc P \vdash \oc P$ starts with the
right rule, as it needs to have everything on the left be of the form $\oc X$.

\begin{mathpar}
  \ebrule{%
    \infer0[Id]{P \vdash P}
    \infer0[Id]{Q \vdash Q}
    \infer2[$\otimes$-r]{P, Q \vdash P \otimes Q}
    \infer1[$\otimes$-l]{P \otimes Q \vdash P \otimes Q}
  }

  \and

  \ebrule{%
    \infer0[Id]{P \vdash P}
    \infer1[Der]{\oc P \vdash P}
    \infer1[Pro]{\oc P \vdash \oc P}
  }
\end{mathpar}