### 2022-06-06 tex/linear.tex

\chapter{Linearity}\label{sec:linearity}

Linear logic was introduced by \citet{girard87linear} as a tool of proof theory.
Its distinguishing feature is that hypotheses, by default, must be used exactly
once.
I devote a chapter to introducing linear logic because it is better understood
than any individual or general semiring-annotated calculi, and is thus an
important point of comparison.
Linear logic will be an instance of the semiring-annotated systems we study
later.

\section{Intuitionistic linear logic}\label{sec:ill}
\input{ill.tex}

\section{DILL}
\input{dill.tex}

\section{Mechanisation survey}
\input{linmech.tex}