### 2022-06-06 tex/simple.tex

\chapter{Mechanisation of simple types}\label{sec:simple}

In this chapter, I review and justify the family of approaches usually used to
represent simple type systems inside dependently typed proof assistants.
These approaches were first presented by \citet{AR99}, who showed a way of
representing well \emph{scoped} terms in a language with polymorphic recursion,
and extended the representation to well \emph{typed} terms in a language with
dependent types.
The representation of terms relies on indexing on both a context --- giving the
types of all the free variables --- and a type for the term itself.
A basic operation on terms is \emph{simultaneous substitution}, which replaces
each variable in the context by a term in another context.
This simultaneous substitution operation fits the form of composition in a
multicategory.
Specifically, in the case of a simple intuitionistic type system, this is the
composition of a Cartesian multicategory, which allows all of the terms being
substituted in to share a context.

\section{Term representation}
\input{terms.tex}
\section{Renaming and substitution}\label{sec:kits}
\input{kits.tex}
\section{Generic semantics}
\input{gen-sem.tex}
\section{Generic syntax}
\section{Natural deduction vs sequent calculus}
\input{ndsc.tex}