### 2022-08-01 tex/wrel.tex

\chapter{Semantics in world-indexed relations}\label{sec:wrel}

While most of the results in this thesis are syntactic in nature, it is
important to make sure that our syntaxes can be interpreted in semantically
interesting ways.
The family of denotational semantics I chose to focus on is those based around
\emph{world-indexed relations}.
Such a semantics allows us to compare runs of a program on two different inputs
(for example allowing us to prove monotonicity with respect to inputs) in a
modal or resource-sensitive way (for example allowing us to prove that a program
is a permutation of its inputs).
The general semantics is intantiated in ways that make use of the particular
semiring used in the syntax for usage restriction.

This chapter can be seen as a refinement and development of the work of
\citet{AbelBernardy2020}.
The subtly different syntax I use with regards to pattern-matching
eliminators~\citep{HMWO21} means that I require fewer and arguably more natural
laws to hold of instantiations of the semantics.
I also provide additional examples to those given by
\citeauthor{AbelBernardy2020}.

\section{Semantics in sets}\label{sec:set-sem}

The eventual semantics based on world-indexed relations will offer theorems
about the behaviour of programs.
The behaviour we care about is just the extension of the program --- what result
we get when semantic values are substituted in for the free variables.
As such, a simple set-based semantics suffices.

The set semantics is the same as one we could give to simply typed
$\lambda$-calculus.
Types are interpreted as sets, with function types interpreted as function sets
and with sum types interpreted as disjoint unions.
Both kinds of products --- with-products and tensor-products --- are interpreted
as Cartesian products, while all $\oc\gr r$ modalities are ignored.
These latter features show us that the set semantics is unaware of usage
restrictions.

% TODO: maybe figure
\begin{displaymath}
  \begin{array}{c@{\hspace{0.5in}}c}
    \begin{eqns}
      \sem{\iota} &=& X_\iota \\
      \sem{\oc\gr rA} &=& \sem{A} \\
      \sem{I} &=& \{*\} \\
      \sem{\top} &=& \{*\} \\
      \sem{0} &=& \{\} \\
    \end{eqns}
    &
    \begin{eqns}
      \\
      \sem{A \multimap B} &=& \sem{A} \rightarrow \sem{B} \\
      \sem{A \otimes B} &=& \sem{A} \times \sem{B} \\
      \sem{A \with B} &=& \sem{A} \times \sem{B} \\
      \sem{A \oplus B} &=& \sem{A} \uplus \sem{B} \\
    \end{eqns}
  \end{array}
\end{displaymath}

Contexts are always interpreted as iterated tensor products of their $\oc$-ed
entries.
In the case of this set semantics, this makes the interpretation of a context
just the Cartesian product of its types, ignoring any usage annotations.
Open terms are interpreted as functions from the interpretation of their context
to the interpretation of their type.

\section{Example results}

The interesting part of the semantics is to prove properties about the set
semantics based on the type of the program.
These ``free theorems'' (in the terminology of \citet{Wadler89}) resemble the
abstraction theorem for parametric polymorphism of \citet{Reynolds83}, but
capture properties the abstraction theorem cannot due to the specificity of
usage-aware typing.
Before discussing technical details, I give some example results below to give a
flavour of what we will be proving.

\paragraph{Linearity}
We should be able to prove, when we instantiate $\name$ with the
$\{\gr0, \gr1, \gr\omega\}$ linearity semiring, that all programs are
permutations, given that they cannot delete or duplicate values freely.
Technically, all programs will be permutations in a designated base type,
because deletion and duplication is derivable for any types made purely of
logical type formers.
To achieve the result, I will assign a ``weight'' to both the context and the
result type of an open term, corresponding to how many values of base type that
term or context needs to support it.
I will then show that all term formers preserve this weight, yielding a
permutation.

\paragraph{Monotonicity}
When we instantiate $\name$ with the $\{\gr{\sim\sim}, \gr{\uparrow\uparrow},
\gr{\downarrow\downarrow}, \gr{\wn\wn}\}$ monotonicity semiring, we expect that
all programs really are monotonic.
Precisely, that means that we impose an ordering on the set denoted by a
designated base type, extend this ordering to all types, then show that passing
(the denotation of) a term bigger inputs makes it produce bigger outputs.
To achieve this, we need to be able to relate two runs of a program.

\paragraph{Sensitivity}
When we instantiate $\name$ with the $\mathbb R^+$ sensitivity semiring, we want
to show that all programs are non-expansive.
Similarly to the case for monotonicity, the notion of distance is semantic,
based on a distinguished base type whose denotation has a notion of distance.

\section{World-indexed relations}

Inspired by modal logic \todo{ref}, our relational semantics assumes a partially
ordered set of worlds $\mathcal W$.
The world keeps track of what resources we have allocated to which parts of the
program.
Our syntax tracks division of usages between subterms, and correspondingly we
have a commutative monoid on $\mathcal W$ to model division of resources.
The monoid structure respects the ordering, i.e., addition of worlds is
monotonic.

{\color{red}intuitively: the elements of $\mathcal W$ determine the
  resources that we are talking about, the pre-order determines a
  notion of ``swappability'', and the monoidal product determines what
  it means for two resources to be combined.}

\begin{example}[Trivial]\label{def:trivial-worlds}
  The one element pre-ordered set $\{*\}$, with ${*} \leq {*}$ and
  ${*} \otimes {*} = {*}$ gives a ``track nothing'' semantics of
  resources. \todo{fwd ref the where this is used.}
\end{example}

\newcommand{\append}{\mathop{++}}

\begin{example}[Lists and Permutations]\label{def:permutation-worlds}
  Let $K$ be some set of values that will be used as keys for sorting.
  I consider multisets over $K$ with finite support, denoted $\mathcal MK$,
  where the multiplicity of each element is either a natural number or $\infty$.
  These multisets represent the keys in scope at any particular point.
  The order on $\mathcal MK$ is multiset equality, plus the fact that
  $\lbag k^\infty \rbag \leq \lbag k^n \rbag$ for any finite $n$ (freely
  extended to respect bag-addition).
  The monoid structure is given by bag-addition.
  The elements with multiplicity $\infty$ exist to have a sound interpretation
  of $\oc\gr\omega$, with $\gr\omega$ coming from the
  $\{\gr0, \gr1, \gr\omega\}$ semiring.
  I am more interested in the elements with finite multiplicity, because in that
  case the ordering corresponds to preservation of keys.

  %Define $\mathcal{P}_K$ to be the pre-ordered set consisting
  %of lists of elements of $K$, with $l_1 \leq l_2$ exactly when $l_1$
  %is a permutation of $l_2$. The monoidal product is defined to be
  %list concatenation: $l_1 \otimes l_2 = l_1 \append l_2$, with the
  %empty list as the monoidal unit.

  %\todo{explain the interpretation of the base type of keys}

  %\todo{explain how this will be used in the $\{0,1,\omega\}$ example.}
\end{example}

\begin{example}[Security Principles]\label{def:privacy-worlds}
  Lattice of sets of names of principles. In terms of resources, this
  can be thought of as the collection of principles that are ``happy''
  with this computation. \todo{still need to fully work out this
    example}
\end{example}

\begin{example}[Distances]\label{def:sensitivity-worlds}
  For sensitivity analysis, I take the partially ordered monoid
  $(\mathbb R^+, \geq, 0, {+})$ of non-negative real numbers under the opposite
  of the usual order.
  We want to establish non-expansiveness of terms, which is why the order is
  reversed.
  The world represents the maximum permissible amount of deviation allowed in
  the result when a unit change is made to the inputs.
\end{example}

\section{Semantics in world-indexed relations}

I give each type former $F$ a semantic counterpart $F^R$.
Where $F$ is an $n$-ary type former, $F^R$ is an operator taking $n$
world-indexed relations and producing another world-indexed relation.
The interpretation of any type is given recursively in the obvious way, with
each base type interpreted as an arbitrary provided world-indexed relation.

\subsubsection{Interpretation of $\multimap$ and $\otimes$}

Given a preorder of semantic resources
$(\mathcal{W}, \leq, \otimes, I)$, we have enough structure to
interpret the function types ($A \multimap B$) and tensor product types
($A \oplus B$) of \name{}.
I define the following notations for connectives of world-indexed relations,
and give interpretations of the type formers in terms of the notations.

\begin{displaymath}
  \arraycolsep=1.5pt
  \begin{array}{rllcl}
    I^R & w & (*,*) &\iff& w \leq I
    \\
    \plr{R \otimes^R S} & w & ((a,b),(a',b'))
    &\iff& \exists x,y.~w = x \otimes y
           \wedge R~x~(a,a')
           \wedge S~y~(b,b')
    \\
    \plr{R \multimap^R S} & w & (f,f')
    &\iff& \forall x.~\forall a,a'.~R~x~(a,a')
           \Rightarrow S~(w \otimes x)~(f~a,f'~a')
  \end{array}
\end{displaymath}

%If we assume we are given semantic interpretations
%$\sem{A}^R : \mathcal{W}^{op} \to \operatorname{Rel}{\sem{A}}$ and
%respectively for $B$, then we define the interpretations
%$\sem{A \multimap B}^R$, $\sem{A \otimes B}^R$, and
%$\sem{I}^R$ as follows:

%\begin{align*}
%  \sem{I}^R &\coloneqq I^R \\
%  \sem{A \otimes B}^R &\coloneqq \sem A^R \otimes^R \sem B^R \\
%  \sem{A \multimap B}^R &\coloneqq \sem A^R \multimap^R \sem B^R
%\end{align*}

The connective $\otimes^R$ is a form of Day convolution~\citep{Day70}.
Using the multiset worlds of \cref{def:permutation-worlds}, $R \otimes^R S$ can
be understood as a pointwise relation on pairs, but where the world (collection
of keys) at which it is true has to be split up into the $R$ world and the $S$
world.
In the implication $R \multimap^R S$, $w$ is the multiset of keys provided by
the function, $x$ is the multiset of keys supporting the argument, and thus all
of the keys in $w \uplus x$ are in the result.
Using the distance worlds of \cref{def:sensitivity-worlds}, we see that the
sensitivity of a tensor product is the sum of the sensitivities of its two
halves, and the sensitivity of the result of an application is similarly the sum
of the sensitivity of the function and the sensitivity of the argument.

\todo{Point out what happens when the monoidal product
  is actually a join}

\subsubsection{Interpretation of lists}

As mentioned in section XXX, we effectively treat lists as tensored
together. \todo{spell out what this entails, and relate it to the
  permutations example}

\subsubsection{Interpretation of $\oc\gr r$}

The choice of interpretation of $\oc\gr r$ links the usage
annotations present in \name{}'s typing contexts to the semantics.

\begin{example}[Trivial]
  The identity
\end{example}

\begin{example}[Linear Types]
  $0$ is the identifies everything and forces the resource to be $I$;
  $1$ is the identity; and $\omega$ forces the resource to be
  duplicable.
\end{example}

\begin{example}[Monotonicity Annotations]
  The worlds we use for monotonicity tracking are trivial, so $\oc^R$ is just
  a transformer of homogeneous binary relations.
  $\oc^R~\gr{\uparrow\uparrow}$ is the identity transformation, while
  $\oc^R~\gr{\downarrow\downarrow}~R$ is $R^{\mathrm{op}}$, the flipped version of
  $R$.
  $\oc^R~\gr{\wn\wn}~R$ is the total relation, while
  $\oc^R~\gr{\sim\sim}~R \coloneqq R \cap R^{\mathrm{op}}$.
\end{example}

\begin{example}[Distances]
  Merp
\end{example}

\subsubsection{Interpretation of the additives: $\with$ and $\oplus$}

Copy the clauses from below.

\begin{displaymath}
  \arraycolsep=1.5pt
  \begin{array}{rllcl}
    \sem{\top}^R & w & (*,*) &\iff& \top \\
    \sem{A \with B}^R & w & ((a,b),(a',b')) &\iff&
    \sem{A}^R~w~(a,a') \wedge \sem{B}^R~w~(b,b') \\
    \sem{A \oplus B}^R & w & (\operatorname{inl} a,\operatorname{inl} a')
                               &\iff& \sem{A}^R~w~(a,a') \\
    \sem{A \oplus B}^R & w & (\operatorname{inr} b,\operatorname{inr} b')
                               &\iff& \sem{B}^R~w~(b,b') \\
  \end{array}
\end{displaymath}

\subsubsection{Assignment of Relations to Types}

Let $\left| X \right|$ be the proposition that the set $X$ is
inhabited.\todo{don't need this if we're just using a preorder.}

\todo{put these in a figure to summarise them, and merge this
  section with the one on the fundamental lemma}

\begin{displaymath}
  \arraycolsep=1.5pt
  \begin{array}{rllcl}
    \sem{\iota}^R & w & (c,c') &\iff& R_\iota~w~(c,c') \\
    \sem{A \multimap B}^R & w & (f,f')
                               &\iff& \forall x,y.~
                                 \left| \mathcal W(x, y \otimes w) \right|
                                 \Rightarrow \\
    &&&& \forall a,a'.~\sem{A}^R~y~(a,a')
                                 \Rightarrow \sem{B}^R~x~(f~a,f'~a') \\
    \sem{\oc\gr rA}^R & w & (a,a') &\iff& \oc\gr r\sem{A}^R~w~(a,a') \\
    \sem{I}^R & w & (*,*) &\iff& \left|\mathcal W(w, I)\right| \\
    \sem{A \otimes B}^R & w & ((a,b),(a',b'))
                               &\iff& \exists x,y.~
                                 \left| \mathcal W(w, x \otimes y) \right|
                                 \wedge \sem{A}^R~x~(a,a')
                                 \wedge \sem{B}^R~y~(b,b') \\
    \sem{\top}^R & w & (*,*) &\iff& \top \\
    \sem{A \with B}^R & w & ((a,b),(a',b')) &\iff&
    \sem{A}^R~w~(a,a') \wedge \sem{B}^R~w~(b,b') \\
    \sem{A \oplus B}^R & w & (\operatorname{inl} a,\operatorname{inl} a')
                               &\iff& \sem{A}^R~w~(a,a') \\
    \sem{A \oplus B}^R & w & (\operatorname{inr} b,\operatorname{inr} b')
                               &\iff& \sem{B}^R~w~(b,b') \\
  \end{array}
\end{displaymath}

A context $\grR\Gamma$ is interpreted as a tensor product of all the
$\oc^R\gr rA$ for each $\gr rA$ in $\grR\Gamma$.
\todo{I use $\grR\gamma$ in other places, but this clashes with
  $\gamma : \sem\Gamma$.}

\begin{displaymath}
 \begin{eqns}
   \sem{\cdot}^R~w~(*,*) &=& \mathcal W(w, I) \\
   \sem{\grR\Gamma, \gr rA}^R~w
   ~((\gamma,a), (\gamma',a')) &=&
   \sem{\grR\Gamma}^R \otimes \oc\gr r\sem A^R
 \end{eqns}
\end{displaymath}

\subsubsection{Semantic Soundness}

We now give the fundamental semantic soundness result for \name{},
that states that the set-theoretic interpretation of every term
(\autoref{sec:set-interp}) preserves the Kripke relational semantics
of contexts and types described above. The proof of this theorem
proceeds by induction on the typing derivation, and has been
formalised in Agda \autoref{sec:agda-formalisation}.

\begin{theorem}
  If $\grR\Gamma \vdash t : T$, then for all worlds $w$ and
  context interpretations $\gamma, \gamma' \in \sem{\Gamma}$, we have:
  \begin{displaymath}
    \sem{\grR\Gamma}^R w~(\gamma, \gamma') \implies \sem{T}^R w~(\sem{t}\gamma, \sem{t}\gamma')
  \end{displaymath}
\end{theorem}

In the very general setting of this section, it is difficult to see
the significance of this result. In the next section, we will
specialise to instantiations of \name{} with specific usage
semirings and specific semantic settings, which will allow us to
derive concrete semantic results of interest.

\section{Examples}