### 2023-01-10 tex/framework.tex
[ writing ] renew generic syntax chapter

\chapter{Generic usage-annotated syntax}\label{sec:framework}

In \cref{sec:semirings}, we saw how to use parametrisation over a partially
ordered semiring to recreate a range of usage-aware calculi.
However, $\name$, with its fixed set of type formers and syntactic forms, is a
long way from capturing the full range of linear-like programming languages
studied in the literature and required in practice.

In this chapter, I take the framework for typed syntaxes with binding developed
by \citet{AACMM21} and apply the principles we discovered in
\cref{sec:semirings} to yield a framework for defining calculi with
semiring-based usage restrictions on variables.
Syntactically, I claim that this framework ranges over all finitiary
variable-based simply typed semiring-annotated calculi, with justification by
comparison to the framework of \citet{AACMM21} and some novel examples in
\cref{sec:other-syntaxes}.
%I also derive analogues to some of the semantic results of \citet{AACMM21},
%strengthening them to take advantage of usage restrictions.

The work in this chapter is fully mechanised in Agda, which allows me to be
precise about the various levels of domain-specific languages which appear.

\input{syntax}

\chapter{Generic usage-aware semantics}\label{sec:semantics}

\input{semantics}

\section{Example semantics}\label{sec:example-semantics}

\input{examples}