### 2023-01-10 tex/syntax.tex

The aim of this chapter is to produce a domain-specific language of
\emph{syntax descriptions}.
One may also use the terminology ``universe'' in place of ``domain-specific
language'', alluding to the \emph{universe pattern} common in dependently typed
programming~\citep{BDJ03}.
A syntax description is essentially a concise, high-level representation of a
type system's syntactic rules.
The information contained in a syntax description is comparable to what is
written (informally) in \cref{fig:lr-bunched}.
The key features allowing these descriptions to capture semiring-annotated
calculi are the distinction between sharing ($\dottimes$) and separating ($*$)
conjunction of premises, modal scaling by a semiring element ($\cdot$), and the
inclusion of semiring annotations on newly bound variables.

%We take the insights of the previous section and use them to build a
%generic framework for posemiring-annotated substructural systems in
%Agda. We will first show \emph{descriptions} of systems, which are
%comprised of rules that have premises combined using the bunched
%combinators. We then show how to construct the Agda data type of
%intrinsically well scoped, typed, and resourced terms for any given
%system of our framework. We use the prototypical system from
%\cref{fig:lr-comb} as a running example. \Cref{sec:other-syntaxes}
%presents further examples that our framework can express.

%We now start to use Agda notation for record and data type
%declarations, to emphasise that our framework has been implemented.

\section{Descriptions of Systems}

% We capture the form of rules exemplified previously\todo{Previously?} via
% \emph{descriptions} of rules.
% The key to making these descriptions work is that they only allow syntactic
% forms that preserve environments.
% These forms are: absence and multiplicity of subterms with the same usage
% annotations, absence and multiplicity of subterms with summed usage annotations,
% scaling of a subterm, and variable binding.\todo{Switching to Agda}

% \paragraph{\AgdaDatatype{Premises}, \AgdaRecord{Rule}s, and \AgdaRecord{System}s.}

I introduce syntax descriptions in three layers: \AgdaRecord{System},
\AgdaRecord{Rule}, and \AgdaRecord{Premises}.
A type \AgdaRecord{System} is made up of multiple \AgdaRecord{Rule}s.
Each \AgdaRecord{Rule} comprises a \AgdaDatatype{Premises} and
a conclusion type. We assume that there is a
$\AgdaBound{Ty} : \AgdaPrimitiveType{Set}$ of types for the system in
scope.

The \AgdaDatatype{Premise} data type describes premises of rules,
using the bunched combinators from \cref{fig:bunched}. A single
premise is introduced by the
\AgdaInductiveConstructor{$\langle$\_`$\vdash$\_$\rangle$}
constructor.  This allows binding of additional variables
\AgdaBound{$\Delta$} (with specified types and usage annotations) and
the specification of a conclusion type \AgdaBound{A} for this premise.
The remaining constructors are descriptions for the
bunched connectives. %, and will be interpreted directly as such, below.

\ExecuteMetaData[\Syntaxtex]{Premises}

A \AgdaRecord{Rule} is a pair of some \AgdaDatatype{Premises} and a
conclusion.
I suggestively use a quoted version of the ``universal entailment'' arrow
$\rightarrowtriangle$, the unquoted version of which interprets the horizontal
line in a traditionally presented typing rule.

\ExecuteMetaData[\Syntaxtex]{Rule}

Finally, a \AgdaRecord{System} consists of a set of rule labels (i.e.,
constructor names), and for each label a description of the
corresponding rule. We use $\rhd$ as infix notation for systems to
associate the label set with the rules.

\ExecuteMetaData[\Syntaxtex]{System}

% \paragraph{\Cref{fig:lr-comb} as a \AgdaRecord{System}.}

As an example, we transcribe a fragment of $\name$ (as defined in
\cref{fig:lr} and \cref{fig:lr-bunched}) into a description.
We give the set of types of
this system as a data type \AgdaDatatype{Ty} (together with a base
type \AgdaInductiveConstructor{$\iota$}). We assume that there is a
posemiring \AgdaInductiveConstructor{Ann} in scope for the
annotations.
There is one label for each instantiation of a logical
rule, but the labels contain no further information about subterms or
restrictions on the context. This will be provided when we associate
labels with \AgdaRecord{Rule}s in a \AgdaRecord{System}.

\noindent
\begin{minipage}[t]{0.5\textwidth}
  \ExecuteMetaData[\PaperExamplestex]{Ty}
  \ExecuteMetaData[\Handtex]{Hand}
\end{minipage}
\begin{minipage}[t]{0.5\textwidth}
  \ExecuteMetaData[\PaperExamplestex]{qlR}
\end{minipage}

To build a system, we associate with each label a rule:

\ExecuteMetaData[\PaperExamplestex]{lR}

Compared to \cref{fig:lr-bunched}, modulo the Agda notation, we can see
that the fundamental structure has been preserved: The rules match
one-to-one, and the bunched premises are the same. A major difference
is that we do not include a counterpart to the
\AgdaInductiveConstructor{var} rule in a
\AgdaRecord{System}. Variables are common to all the systems
representable in our framework.

\section{Terms of a \AgdaRecord{System}}\label{sec:terms}

The next thing we want to do is to build terms in the described type system.
The following definitions are useful for talking about types indexed over
contexts, judgement forms, and judgement forms admitting newly bound variables,
respectively.

\ExecuteMetaData[\Syntaxtex]{OpenFam}

To specify the meaning of descriptions, we assume some \AgdaBound{X} : \AgdaFunction{ExtOpenFam},
% \ExecuteMetaData[\Interpretationtex]{X},
over which we form one layer of syntax, using the function
\AgdaFunction{$\llbracket$\_$\rrbracket$p} that interprets
\AgdaDatatype{Premises} defined below.  The first argument to
\AgdaBound{X} is the new variables bound by this layer of syntax, as
exemplified in the first clause of
\AgdaFunction{$\llbracket$\_$\rrbracket$p}.  The second argument is
the context containing the variables being carried over from the
previous layer.  Notice that this is not, in general, the same as the
context from the previous layer, because the usage annotations may
have been changed by connectives like
\AgdaInductiveConstructor{\_`$*$\_} and
\AgdaInductiveConstructor{\_`$\cdot$\_}.  The third argument is the
type of subterm required.

The remainder of the clauses of \AgdaFunction{$\llbracket$\_$\rrbracket$p}
are the interpretation into bunched combinators.
The superscript \AgdaFunction{$^c$} on the bunched connectives denotes that
they have been lifted from predicates on usage vectors to predicates on
contexts, with the type component of the context shared throughout.
Additive connectives \AgdaFunction{$\dot1$} and \AgdaFunction{$\dottimes$} are
already polymorphic (not relying on anything specific about usage vectors), so
do not need a \AgdaFunction{$^c$} variant.

\ExecuteMetaData[\Interpretationtex]{semp}

The interpretation of a \AgdaRecord{Rule} checks that the rule targets
the desired type and then interprets the rule's premises
\AgdaBound{ps}.  Notice that the interpretation of the premises is
independent of the conclusion of the rule, which accounts for the use
of \AgdaFunction{OpenType} in
\AgdaFunction{$\llbracket$\_$\rrbracket$p} versus
\AgdaFunction{OpenFam} in \AgdaFunction{$\llbracket$\_$\rrbracket$r}.

\ExecuteMetaData[\Interpretationtex]{semr}

The interpretation of a \AgdaRecord{System} is to choose a rule label
\AgdaBound{l} from \AgdaBound{L} and interpret the corresponding rule
\AgdaBound{rs}\AgdaSpace{}\AgdaBound{l} in the same context and for the same
conclusion.

\ExecuteMetaData[\Interpretationtex]{sems}

The most obvious way to make such an \AgdaBound{X} is to use some existing
\AgdaFunction{OpenFam} on an extended context.
We defined \AgdaFunction{Scope} to do this: take the new variables
\AgdaBound{$\Delta$}, concatenate them onto the existing context
\AgdaBound{$\Gamma$}, and pass the extended context onto the judgement
\AgdaBound{T}.

\ExecuteMetaData[\Syntaxtex]{Scope}

%{\color{red}(Forward ref: for now, we could have inlined \texttt{Scope}.)}

We use \AgdaFunction{Scope} to deal with new variables in syntax.
Terms resemble the free monad over a layer-of-syntax functor, though
that picture is complicated by variable binding.  A term is either a
variable or a use of a logical rule together with terms for each of
the required subterms. The \AgdaFunction{Size} argument is a use of
Agda's sized types to record that subterms are smaller than the
surrounding term for the termination checker.

\ExecuteMetaData[\Termtex]{Term}

This definition uses \AgdaFunction{$\dotto$}, which, analogously to
\AgdaFunction{$\dottimes$}, is an index-preserving version of the function
space.
We take \AgdaFunction{$\dotto$} to handle $n$ many indices --- in this
case two (the context and the type).
The notation
\AgdaFunction{$\forall[$}\AgdaSpace{}\AgdaBound{T}\AgdaSpace{}\AgdaFunction{]}
stands for
\AgdaSymbol{$\forall$}\AgdaSpace{}\AgdaSymbol{\{}%
\AgdaBound{x$_1$}\AgdaSpace{}\AgdaSymbol{$\ldots$}\AgdaSpace{}\AgdaBound{x$_n$}%
\AgdaSymbol{\}}\AgdaSpace{}\AgdaSymbol{$\to$}\AgdaSpace{}\AgdaBound{T}%
\AgdaSpace{}%
\AgdaBound{x$_1$}\AgdaSpace{}\AgdaSymbol{$\ldots$}\AgdaSpace{}\AgdaBound{x$_n$},
where \AgdaBound{T} is a type family with $n$ many indices.

Terms in this data type are difficult to write by hand, due to the
need for proofs that the usage contexts are handled correctly. For
example, the following term is needed to show that, in the $\{\gr0,
\gr1, \gr\omega\}$ (linearity) posemiring of \cref{def:lin-semiring},
$\oc\gr\omega$ forms a comonad.
Pattern synonyms \AgdaInductiveConstructor{$\multimap$I},
\AgdaInductiveConstructor{!E$'$}, and
\AgdaInductiveConstructor{!I$'$} stand for applications of
\AgdaInductiveConstructor{`con}, with the latter two taking explicit usage
contexts and proofs.
On concrete
posemirings (as in this example), unification is particularly poor at
inferring the usage contexts from the proofs because addition and
multiplication are no longer (judgementally) injective.
The function \AgdaFunction{var\#} is a way of turning a statically known de
Bruijn level and a usage proof into an application of \AgdaInductiveConstructor{`var}.

\ExecuteMetaData[\HeavyItex]{cojoin-explicit}

 Writing terms like this
is clearly unsustainable. We will see a way of automating the
necessary proofs via a \AgdaRecord{System}-generic elaborator in
\cref{sec:usage-elaborator}.

%Here is an example term, using the \AgdaFunction{$\lambda$R} system.
%First, for ease of writing, we introduce pattern synonyms for each of the
%typing rules we use.

%\ExecuteMetaData[\PaperExamplestex]{patterns}

%Our example term is a function that flips a tagged union wrapped in an
%arbitrarily annotated \emph{bang}.
%Much of the effort in writing such a term goes into writing the various
%relatedness proofs between usage contexts --- observing, for example, that two
%usage contexts sum together to make a third, or that a usage context used for
%a variable is a basis vector.
%We give a method of automating these proofs in \cref{sec:usage-elaborator}.
%\todo{To be clear, we don't actually write this.}

%\ExecuteMetaData[\HeavyItex]{lR-term}

% A layer of syntax supports the following functorial action.

% \ExecuteMetaData[\Maptex]{map-s-type}

\section{More example syntaxes}\label{sec:other-syntaxes}

With the range of representable syntaxes now formalised, we can explore encoding
techniques for syntaxes more exotic than ST$\lambda$C and $\name$.
As well as the variations presented in \cref{sec:variant}, we can represent a
usage-annotated $\mu\tilde\mu$-calculus and a Linear/non Linear system.

\subsection{The system $\mu\tilde\mu$}
I encode a usage-annotated version of System $L$/the
$\mu\tilde\mu$-calculus~\citep{CH00} --- a syntax for classical logic --- in
such a way that contexts capture the undistinguished parts of the sequent.
As such, the generic substitution lemma we get in \cref{sec:kit-to-sem} is the
form of substitution required in standard $\mu\tilde\mu$-calculus metatheory.
Though the $\mu\tilde\mu$-calculus is originally described as a sequent
calculus~\citep{CH00}, I use the techniques of
\citet[p.~12]{herbelin-hab} and \citet{LC06} to present it using hypothetical
judgements, thus giving a notion of \emph{variable} to the system.

Unlike the single judgement form of \name{} and standard simply typed
$\lambda$-calculi, the $\mu\tilde\mu$-calculus has three judgement forms:
terms, coterms, and commands.
Read logically, terms and coterms are seen to, respectively, prove and refute
propositions (types), while commands exhibit contradictions.
This means that the abstract \AgdaBound{Ty} in the generic framework is
instantiated to \AgdaDatatype{Conc} (for \emph{conclusion}) as below, with
\AgdaDatatype{Ty} not being exposed directly to the generic framework.
For now, I just consider multiplicative disjunction $\parr$ (\emph{par}) and
negation/duality, beside an uninterpreted base type.
These are enough to exhibit classical behaviour.

\noindent
\begin{minipage}[t]{0.5\textwidth}
  \ExecuteMetaData[\MuMuTildetex]{Ty}
\end{minipage}
\begin{minipage}[t]{0.5\textwidth}
  \ExecuteMetaData[\MuMuTildetex]{Conc}
\end{minipage}

With \AgdaBound{Ty} instantiated as \AgdaDatatype{Conc}, all terms are assigned
\AgdaDatatype{Conc} type, as are all the variables.
No variables are given \AgdaInductiveConstructor{com} type, similar to how in
the bidirectional typing syntax of \citet[p.~25]{AACMM21}, no variables are
given \AgdaInductiveConstructor{Check} type.
How to observe this invariant is covered in the latter paper, so we will not
repeat it here (having not yet seen how to write traversals on terms).

The syntax comprises a \emph{cut} between a term and a coterm of the same type,
the eponymous $\mu$ and $\tilde\mu$ constructs for proof by contradiction, and
then term and coterm (introduction and elimination) forms for negation and
\emph{par}.

\ExecuteMetaData[\MuMuTildetex]{MMT}

%With a collection of pattern synonyms and the machinery from
%\cref{sec:usage-elaborator}, we can write an example term: a function which
%flips the disjuncts of a \emph{par}.

%\ExecuteMetaData[\MuMuTildeTermtex]{patterns}
%\ExecuteMetaData[\MuMuTildeTermtex]{myComm}

\subsection{Duplicability and L/nL}\label{sec:dup-lnl}
In \cref{sec:rec}, I introduced a bunched connective $\Box^{0{+}}$, used to
ensure that such a premise is derived in a context that allows it to be
duplicated and discarded.
Syntax for this connective can be added to \AgdaDatatype{Premises} and worked
through the various parts of the framework, though I do not show these parts in
the text of this thesis for the sake of simplicity and brevity.
Additionally, variations such as $\Box^0$ (requiring only that the context is
${} \leq \gr0$) and $\Box^{0{+}{*}}$ (requiring, in addition to
$\Box^{0{+}}$, closure under multiplication by any scalar) can be added to the
framework independently of each other and of $\Box^{0{+}}$.

As well as the use case of inductive types, as shown in \cref{sec:rec}, I use
the $\Box^{0{+}{*}}$-modality to encode the restriction to intuitionistic
variables in the intuitionistic judgement forms of the L/nL
calculus~\citep{Benton94}.
I will explain this in detail in due course.

\todo{Maybe partially merge with the later L/nL section.}
Linear/non-Linear Logic has two sorts of types: linear types $A, B, C$ and
intuitionistic types $X, Y, Z$.
I encode these two sorts in the Agda types
\AgdaDatatype{Ty}\AgdaSpace{}\AgdaInductiveConstructor{lin} and
\AgdaDatatype{Ty}\AgdaSpace{}\AgdaInductiveConstructor{int}, respectively.
Using an indexed type lets me take the total space of \AgdaDatatype{Ty}, named
\AgdaFunction{$\Sigma$Ty}, as the evident dependent pair type.
\AgdaFunction{$\Sigma$Ty} becomes the type of types seen by the framework.
The types themselves are, on the linear side,
a tensor unit \AgdaInductiveConstructor{tI},
a tensor product \AgdaInductiveConstructor{\_t$\otimes$\_}, and
a linear function space \AgdaInductiveConstructor{\_t$\multimap$\_};
on the intuitionistic side,
a unit \AgdaInductiveConstructor{t1},
a product \AgdaInductiveConstructor{\_t$\times$\_}, and
an intuitionistic function space \AgdaInductiveConstructor{\_t$\to$\_};
and adjoint type formers \AgdaInductiveConstructor{tF} and
\AgdaInductiveConstructor{tG}.
I also throw in a linear base type \AgdaInductiveConstructor{$\iota$}.

\ExecuteMetaData[\LnLtex]{Frag}
\ExecuteMetaData[\LnLtex]{Ty}
%\ExecuteMetaData[\LnLtex]{SigTy}

As in the other examples, I define a data type of rule labels
\AgdaDatatype{`LnL} before defining the type system \AgdaFunction{LnL}.
The system \AgdaFunction{LnL} has types \AgdaFunction{$\Sigma$Ty} and usage
annotations $\{\gr0, \gr1, \gr\omega\}$, as we have used before for linear
systems.

We want to be able to embed the L/nL judgement forms
$\Theta; \Gamma \vdash_{\mathcal L} A$ and $\Theta \vdash_{\mathcal C} X$, where
$\Theta$ is made of intuitionistic variables and $\Gamma$ is made of linear
variables.
Following allowed usage patterns, intuitionistic variables will all have usage
annotation $\gr\omega$, while linear variables will each be annotated
either $\gr1$ or $\gr0$.
These assignments of usage annotations are enforced wherever the rules bind new
variables.
By usage subsumption, intuitionistic variables may also end up with annotation
$\gr1$ or $\gr0$, but we see this as a (meaningless) refinement of the
$\gr\omega$ annotation, not causing a soundness problem.
The key invariant is that linear variables never get annotation $\gr\omega$.

The framework already tells us what variables and contexts we can
have --- namely that we can make any assumption in any context.
This means that we can propose ill formed sequents like
$\gr1A \vdash X$.
To make sure that this is not a problem, we want to maintain the invariant that
any \emph{derivation} of a well formed sequent contains only well formed
sequents.
I enforce this locally using the $\Box^{0{+}{*}}$ bunched modality:
Every rule targeting an intuitionistic judgement has its premises wrapped in a
$\Box^{0{+}{*}}$ to restrict the usage annotations to $\gr0$ and $\gr\omega$.
Additionally, rules \TirName{$F$i} and \TirName{$G$e}, which target linear
judgements but have intuitionistic premises, similarly get a $\Box^{0{+}{*}}$.
Note that having the annotation $\gr0$ means that well formed intuitionistic
sequents can contain linear variables annotated $\gr0$, but these variables are
not usable.

The rules are presented in Agda code below, and also using bunched inference
rule notation in \cref{fig:LnL-bunched}

\ExecuteMetaData[\LnLtex]{quoteLnL}
\ExecuteMetaData[\LnLtex]{LnL}

\begin{figure}
  \begin{mathpar}
    \ebrule[comb]{%
      \hypo{I^{\sep}}
      \infer1[$I$i]{\vdashL I}
    }
    \and
    \ebrule[comb]{%
      \hypo{\vdashL I}
      \hypo{\sep}
      \hypo{\vdashL C}
      \infer3[$I$e]{\vdashL C}
    }
    \and
    \ebrule[comb]{%
      \hypo{\vdashL A}
      \hypo{\sep}
      \hypo{\vdashL B}
      \infer3[$\otimes$i]{\vdashL A \otimes B}
    }
    \and
    \ebrule[comb]{%
      \hypo{\vdashL A \otimes B}
      \hypo{\sep}
      \hypo{\gr1A, \gr1B \vdashL C}
      \infer3[$\otimes$e]{\vdashL C}
    }
    \and
    \ebrule[comb]{%
      \hypo{\gr1A \vdashL B}
      \infer1[$\multimap$i]{\vdashL A \multimap B}
    }
    \and
    \ebrule[comb]{%
      \hypo{\vdashL A \multimap B}
      \hypo{\sep}
      \hypo{\vdashL A}
      \infer3[$\multimap$e]{\vdashL B}
    }
    \and
    \ebrule[comb]{%
      \hypo{\Boxzpt\plr{\vdashC X}}
      \infer1[$F$i]{\vdashL FX}
    }
    \and
    \ebrule[comb]{%
      \hypo{\vdashL FX}
      \hypo{\sep}
      \hypo{\gr\omega X \vdashL C}
      \infer3[$F$e]{\vdashL C}
    }
    \and
    \ebrule[comb]{%
      \hypo{\Boxzpt\dot1}
      \infer1[$1$i]{\vdashC 1}
    }
    \and
    \text{(no \TirName{$1$e})}
    \and
    \ebrule[comb]{%
      \hypo{\Boxzpt(\vdashC X}
      \hypo{\dottimes}
      \hypo{\vdashC Y)}
      \infer3[$\times$i]{\vdashC X \times Y}
    }
  \end{mathpar}
  \caption{Linear/non-Linear Logic in bunched notation}\label{fig:LnL-bunched}
\end{figure}

Later,\todo{When?} I will give translations between this encoding of L/nL and
$\name$, providing some assurance that the correct logic has been encoded.

%There is one more bunched combinator we have experimented with adding to the
%framework:
%
%\[
%  \plr{\Box T}\,\grR \coloneqq \Sigma\grRprime.~\plr{\grRprime \leq \grR}
%  \times \plr{\grRprime \leq \gr0}
%  \times \plr{\grRprime \leq \grRprime + \grRprime}
%  \times T\,\grRprime
%\]
%
%The idea of $\plr{\Box T}\,\grR$ is to assert that $\grR$, or some refinement
%of it, can be both discarded and duplicated indefinitely, and in the
%refinement we have a $T$.
%We use this combinator to introduce subterms that are used an unknown number of
%times, for example the continuations of the eliminator of an inductive type,
%or other fixed points.
%We can also use it in linear/non-linear style systems~\cite{Benton94} to make
%sure linear variables are not available in the intuitionistic fragment.
%
%Adding the $\Box$ combinator is the only thing we have found that requires our
%linear maps be functional rather than merely relational.

\section{A layer of syntax is functorial}\label{sec:functorial}

A basic property of the universe of syntaxes
is that every syntax supports a functorial action on subterms, realised by a
function \AgdaFunction{map-s}.
Its type says that to map a function \AgdaBound{f}
over a layer of syntax, there must be a linear map \AgdaBound{$\Psi$} relating the
domain and codomain usage contexts, and \AgdaBound{f} should be usable
wherever the domain and codomain usage contexts are similarly related by
\AgdaBound{$\Psi$}.
The notation
\AgdaBound{$\Psi$}\AgdaSpace{}\AgdaSymbol{.}\AgdaField{rel}\AgdaSpace{}%
\AgdaBound{P}\AgdaSpace{}\AgdaBound{Q}
means the same as $\grQ \leq \grP\gr\Psi$ from \cref{sec:ren-sub-lr}.
The new notation comes about because I state \AgdaBound{$\Psi$} as a total and
deterministic linear relation, rather than a linear map, for reasons related to
mechanisation beyond the scope of this thesis.

\ExecuteMetaData[\Maptex]{map-s-type}

This generality is needed because usage contexts change between
a term and its immediate subterms---they are decomposed according to the bunched connectives used in the rules.
\AgdaBound{X} and \AgdaBound{Y} are \AgdaFunction{ExtOpenFam}s, with
\AgdaBound{$\Theta$} being the context extension for a subterm (i.e., the
variables newly bound in that subterm).
Unlike usage annotations, types in the contexts \AgdaBound{$\gamma$} and \AgdaBound{$\delta$}, and the conclusion types implicit here, are preserved throughout.
This is the essence of the usage annotation based approach---we use traditional techniques for variable binding, with the usage annotations layered on top.

The heart of \AgdaFunction{map-s} is \AgdaFunction{map-p}, which recursively
works through the structure \AgdaBound{ps} of premises of the rule applied,
acting on each subterm it finds.
Here, particularly in the clauses for \AgdaInductiveConstructor{`$\sep$} and
\AgdaInductiveConstructor{`$\cdot$}, we see why it is not enough for the
function on subterms to apply at usage contexts \AgdaBound{P} and \AgdaBound{Q}
--- rather, it also needs to apply at any similarly related \AgdaBound{P$'$}
and \AgdaBound{Q$'$}.
In the case of \AgdaInductiveConstructor{`$\sep$}, we have that
$\grP \leq \grP_M + \grP_N$, with \AgdaBound{M} and \AgdaBound{N} being
collections of subterms in usage contexts $\grP_M$ and $\grP_N$, respectively.
Linearity of \AgdaBound{$\Psi$} yields $\grQ_M$ and $\grQ_N$ such that
$\grQ \leq \grQ_M + \grQ_N$ and we use \AgdaFunction{map-p} recursively at
$(\grP_M, \grQ_M)$ and $(\grP_N, \grQ_N)$ on \AgdaBound{M} and \AgdaBound{N}.
The cases for \AgdaInductiveConstructor{`$\cdot$} and
\AgdaInductiveConstructor{`$I^*$} are similar, each using a different aspect
of linearity.
In contrast, the cases for \AgdaInductiveConstructor{`$\dot1$} and
\AgdaInductiveConstructor{`$\dot\times$}, which are the only constructors used in fully structural
systems, do not involve any changes in the usage contexts.

The linearity of relation \AgdaBound{$\Psi$} is given by fields
\AgdaField{rel-0$_m$}, \AgdaField{rel-+$_m$}, and \AgdaField{rel-*$_m$} (with
the subscript-m being a mnemonic for \emph{module}, as opposed to scalar).
\todo{Refer to preliminaries}

\ExecuteMetaData[\Maptex]{map-p}

I have also extended \AgdaFunction{map-p} to handle the various
$\Box$-modalities described in \cref{sec:dup-lnl}.
The Agda code for this extension is not particularly readable, so I do not
include it in this document.
However, this extension is notable as the only part of the framework requiring
that the linear relation \AgdaBound{$\Psi$} be functional (i.e., total and
deterministic).