### 2023-02-01 tex/rec.tex

The bunched connectives give us a means by which to quickly design and
experiment with new type systems.
We will see in \cref{sec:framework} that all such syntaxes are well behaved to
the extend that they all support the appropriate notion of substitution, which
again expediates experimentation.
In this section, I give some example syntactic features that can be covered in
the bunched connective paradigm.

\subsection{Alternative object-language connectives}\label{sec:alt}

I presented $\name$ as an essentially arbitrary collection of rules, with no
broader characterisation than that it intuitively seems to capture some notion
of ``usage'' and that it supports the semantics I give in \cref{sec:wrel}.
Therefore, in this subsection I give some examples of variant syntaxes others,
with different use cases, may be interested in.
In particular, I give a variant of function types where usage annotations go
``on the arrow'', and a variant of tensor products with a stronger elimination
rule.

We can produce an annotated function arrow connective by combining parts of the
rules for the unannotated function arrow $\multimap$ with parts of the rules for
the modal operator $\oc$.

\begin{mathpar}
  \ebrule[comb]{%
    \hypo{\gr rA \vdash B}
    \infer1[$\gr r\multimap$-I]{\vdash \gr rA \multimap B}
  }
  \and
  \ebrule[comb]{%
    \hypo{\vdash \gr rA \multimap B}
    \hypo{\sep}
    \hypo{\gr r \cdot \plr{\vdash A}}
    \infer3[$\gr r\multimap$-E]{\vdash B}
  }
\end{mathpar}

We can also give an alternative elimination rule for $\otimes$-products, as
found in Granule~\citep{Granule18,HMWO21} and the work of
\citet{AbelBernardy2020} and \citet{reed10distance}.
This modified rule allows us to, for example, have a variable
$\gr rx : A \otimes B$ for any $\gr r$, and pattern-match it into
$\gr ry : A, \gr rz : B$.
With the standard $\name$ rule from \cref{fig:lr}, the newly bound variables are
always given annotation $\gr 1$, and we would only be able to do the match in
the first place if $\gr r$ were coercible to (i.e.\ less than or equal to)
$\gr1$.
The alternative rule below is not compatible with linear logic, because it
allows us to derive
$\oc\gr r\plr{A \otimes B} \multimap \oc\gr rA \otimes \oc\gr r B$
parametric in $\gr r$, $A$, and $B$.

\begin{mathpar}
  \ebrule[comb]{%
    \hypo{\gr r \cdot \plr{\vdash A \otimes B}}
    \hypo{\sep}
    \hypo{\gr rA, \gr rB \vdash C}
    \infer3[$\otimes$-E$'$]{\vdash C}
  }
\end{mathpar}

\subsection{Adding inductive types and recursion}\label{sec:rec}

Based on an intuitive understanding of ``usage'', recursion introduces a new
phenomenon relative to the forms of programs we have seen so far:
Terms can be used an unbounded number of times.
For example, notice the following reduction in Agda.

\ExecuteMetaData[\SnippetsTwotex]{sum-reduction}

The function \AgdaFunction{\_+\_} has been copied into 3 different places in
the running of the program.
This copying is despite no type telling us that \AgdaFunction{\_+\_} would be
used 3 times (both \verb|[1,2,3]| and \verb|[2,3]| have type
\AgdaDatatype{List}\AgdaSpace{}\AgdaDatatype{$\mathbb N$}, despite the
corresponding folds using \AgdaFunction{\_+\_} a different number of times).
As such, when checking an application of \AgdaFunction{foldr}, we need check
that we can use its functional argument (\AgdaFunction{\_+\_} in this case) an
arbitrary number of times.
If we were to fix $\Ann$ as the $\{\gr0, \gr1, \gr\omega\}$ posemiring, then
wrapping the type of the functional argument in $\oc\gr\omega$ would suffice.
However, we want to remain generic in the choice of semiring.

The following additions to \name{} support a broad class of
inductive types.
I define strictly positive functors syntactically, with the only notable
restrictions being not being allowed to use the type variable $X$ in the domain
of a function type and within a $\oc$-type.
I then add least fixed points of such strictly positive functors to the syntax
of types.

\begin{align*}
  U &\Coloneqq A \multimap (-) \\
  {\odot} &\Coloneqq {\otimes} \mid {\oplus} \mid {\with} \\
  F[X], G[X] &\Coloneqq X \mid A \mid U(F[X]) \mid F[X] \odot G[X] \\
  A &\Coloneqq \cdots \mid \mu X.~F[X]
\end{align*}

\begin{example}
  We may define $\mathrm{List}_A \coloneqq \mu X.~I \oplus \plr{A \otimes X}$.
\end{example}

In the typing rules, introduction of an inductive type is standard.
For the elimination rule, we follow a similar pattern to other pattern-matching
rules --- $\oplus$-E, $\otimes$-E, and $\oc$-E --- by splitting the context
and typing the eliminand in one half ($\grP$).
We type the continuation in the other half, but because the continuation may
be used multiple times, and in a modal context, we require that $\grQ$ is
preserved by all linear operations.

\begin{displaymath}
  \begin{prooftree}
    \hypo{\grR\gamma \vdash F[\mu X.~F[X]]}
    \infer1[$\mu$-I]{\grR\gamma \vdash \mu X.~F[X]}
  \end{prooftree}
\end{displaymath}
\begin{displaymath}
  \begin{prooftree}
    \hypo{\grR \leq \grP + \grQ}
    \hypo{\grP\gamma \vdash \mu X.~F[X]}
    \hypo{%
      \begin{matrix*}[l]
        \grQ \leq \gr0 \\
        \grQ \leq \grQ + \grQ
      \end{matrix*}%
    }
    \hypo{\grQ\gamma, \gr1F[C] \vdash C}
    \infer4[$\mu$-E]{\grR\gamma \vdash C}
  \end{prooftree}
\end{displaymath}

\begin{example}\label{thm:list-rules}
  For lists, we can derive the following introduction and elimination rules
  (with usage constraints in the application of \TirName{$\mu$-E} in the
  \TirName{foldr} rule, relating $\grR$ to $\grP$ and $\grQ$ and restricting
  $\grQ$, omitted to save space).

  \begin{align*}
    \TirName{nil}\left[
    \begin{prooftree}
      \hypo{\grR \leq \gr0}
      \infer1[$I$-I]{\grR\gamma \vdash I}
      \infer1[$\oplus$-I$_0$]%
      {\grR\gamma \vdash I \oplus \plr{A \otimes \mathrm{List}_A}}
      \infer1[$\mu$-I]{\grR\gamma \vdash \mathrm{List}_A}
    \end{prooftree}
    \right.
    &&
    \TirName{cons}\left[
    \begin{prooftree}
      \hypo{\grR \leq \grP + \grQ}
      \hypo{\grP\gamma \vdash A}
      \hypo{\grQ\gamma \vdash \mathrm{List}_A}
      \infer3[$\otimes$-I]{\grR\gamma \vdash A \otimes \mathrm{List}_A}
      \infer1[$\oplus$-I$_1$]%
      {\grR\gamma \vdash I \oplus \plr{A \otimes \mathrm{List}_A}}
      \infer1[$\mu$-I]{\grR\gamma \vdash \mathrm{List}_A}
    \end{prooftree}
    \right.
  \end{align*}
  \begin{displaymath}
    \TirName{foldr}\left[
    \begin{prooftree}
      \hypo{\grP\gamma \vdash \mathrm{List}_A}
      \infer0[Var]{\gr0\gamma, \gr1\plr{I \oplus \plr{A \otimes C}}
        \vdash I \oplus \plr{A \otimes C}}
      \hypo{\nabla^n}
      \hypo{\nabla^c}
      \infer3[$\oplus$-E]{\grQ\gamma, \gr1\plr{I \oplus \plr{A \otimes C}}
        \vdash C}
      \infer2[$\mu$-E]{\grR\gamma \vdash C}
    \end{prooftree}
    \right.
  \end{displaymath}
  \begin{align*}
    \textrm{where }\nabla^n &\coloneqq
    \begin{prooftree}
      \infer0[Var]{\gr0\gamma, \gr1I \vdash I}
      \hypo{\grQ\gamma \vdash C}
      \infer1[Wk]{\grQ\gamma, \gr0I \vdash C}
      \infer2[$I$-E]{\grQ\gamma, \gr1I \vdash C}
      \infer1[Wk]{\grQ\gamma, \gr0\plr{I \oplus \plr{A \otimes C}}, \gr1I
        \vdash C}
    \end{prooftree}
    \\\\
    \textrm{and }\nabla^c &\coloneqq
    \begin{prooftree}
      \infer0[Var]{\gr0\gamma, \gr1\plr{A \otimes C} \vdash A \otimes C}
      \hypo{\grQ\gamma, \gr1A, \gr1C \vdash C}
      \infer1[Wk]{\grQ\gamma, \gr0\plr{A \otimes C}, \gr1A, \gr1C \vdash C}
      \infer2[$\otimes$-E]{\grQ\gamma, \gr1\plr{A \otimes C} \vdash C}
      \infer1[Wk]%
      {\grQ\gamma, \gr0\plr{I \oplus \plr{A \otimes C}}, \gr1\plr{A \otimes C}
        \vdash C}
    \end{prooftree}
  \end{align*}
\end{example}

Following \cref{sec:lnd}, I want to turn the ad hoc constraints on $\grP$,
$\grQ$, and $\grR$ into the result of some premise combinators.
To do this, I introduce a new combinator $\Box^{0{+}}$ defined below,
along with the resulting implicit-context typing rules.

\ExecuteMetaData[\Bunchedtex]{Box}

\begin{align*}
  \begin{prooftree}[comb]
    \hypo{\vdash F[\mu X.~F[X]]}
    \infer1[$\mu$-I]{\vdash \mu X.~F[X]}
  \end{prooftree}
  &&
  \begin{prooftree}[comb]
    \hypo{\vdash \mu X.~F[X]}
    \hypo{\sep}
    \hypo{\Box^{0{+}}\plr{\gr1F[C] \vdash C}}
    \infer3[$\mu$-E]{\vdash C}
  \end{prooftree}
\end{align*}

\begin{example}
  We can state the rules for lists derived in \cref{thm:list-rules} as follows.
  \begin{align*}
    \begin{prooftree}[comb]
      \hypo{I^*}
      \infer1{\vdash \mathrm{List}_A}
    \end{prooftree}
    &&
    \begin{prooftree}[comb]
      \hypo{\vdash A}
      \hypo{\sep}
      \hypo{\vdash \mathrm{List}_A}
      \infer3{\vdash \mathrm{List}_A}
    \end{prooftree}
    &&
    \begin{prooftree}[comb]
      \hypo{\vdash \mathrm{List}_A}
      \hypo{\sep}
      \hypo{\Box^{0{+}}
        \plr{\vdash C\hskip0.75em\dottimes\hskip0.75em\gr1A, \gr1C \vdash C}}
      \infer3{\vdash C}
    \end{prooftree}
  \end{align*}
\end{example}

Among the inductive types given by this $\mu$ operator are also
\emph{$\with$-lists} (\emph{with-lists}).

\begin{example}
  Let us define \emph{$\with$-lists} as follows.
  \[
    \mathrm{WithList}_A \coloneqq \mu X.~\top \oplus \plr{A \with X}
  \]
  Unlike ordinary lists (which we may call \emph{$\otimes$-lists}),
  $\with$-lists support a \emph{lookup} operation with the type below, where
  $\mathbb N \coloneqq \mu X.~I \oplus X$.
  \begin{align*}
    \mathit{lookup} :
    \mathrm{WithList}_A \multimap \mathbb N \multimap \top \oplus A
  \end{align*}
  The $\name$ term implementing $\mathit{lookup}$ in terms of iterators
  (\TirName{$\mu$-E}) is complex and tedious, so I do not show it here.
  Instead, I provide the following clausal specification, where
  $\mathrm{nil}\plr{-}$ and $\mathrm{cons}\plr{-}$ are the constructors of
  $\with$-lists, $\mathrm{zero}\plr{}$ and $\mathrm{succ}\plr{}$ are the
  constructors of $\mathbb N$, $\mathrm{inl}\plr{-}$ and $\mathrm{inr}\plr{-}$
  are the constructors of $\oplus$, $\langle\rangle$ constructs the
  inhabitant of $\top$, and postfix $.\pi_l$ and $.\pi_r$ are the projections
  out of a $\with$-pair.
  %\begin{align*}
  %  \mathit{lookup}\,\mathrm{nil}\plr{t}\,i
  %  &\coloneqq \mathrm{inl}\plr{\langle\rangle} \\
  %  \mathit{lookup}\,\mathrm{cons}\plr{\mathit{xxs}}\,\mathrm{zero}\plr{}
  %  &\coloneqq \mathrm{inr}\plr{\mathit{xxs}.\pi_l} \\
  %  \mathit{lookup}\,\mathrm{cons}\plr{\mathit{xxs}}\,\mathrm{succ}\plr{i}
  %  &\coloneqq \mathit{lookup}\,\plr{\mathit{xxs}.\pi_r}\,i
  %\end{align*}

  %\noindent{}\hspace{-0.75ex}
  \[
    \begin{pmboxed}
      \defaultcolumn{l}
      \> \mathit{lookup} \> \mathrm{nil}\plr{t} \> i
      \> \coloneqq \> \mathrm{inl}\plr{\langle\rangle} \\
      \> \mathit{lookup} \> \mathrm{cons}\plr{\mathit{xxs}}
      \> \mathrm{zero}\plr{} \> \coloneqq
      \> \mathrm{inr}\plr{\mathit{xxs}.\pi_l} \\
      \> \mathit{lookup} \> \mathrm{cons}\plr{\mathit{xxs}}
      \> \mathrm{succ}\plr{i} \> \coloneqq
      \> \mathit{lookup}\,\plr{\mathit{xxs}.\pi_r}\,i \\
    \end{pmboxed}
  \]
  %\begin{verbatim}
  %  lookup := \xs. wfoldr(xs;
  %    List I -o T + A;
  %    t. \i. foldr(i; T -o T + A; \t'. inl(t'); u, r. let () := u in r) t;
  %    xr. \i. let (i', r) :=
  %      foldr(i;
  %        A & (T + A) -o List I * (T + A);
  %        \xr'. ([], prl(xr'));
  %        u, ir. \xr'.
  %          let () := u in
  %          let (i', r) := ir in
  %          (() :: i', prr(xr') prl(ir)))
  %      xr in
  %      foldr(i'; T + A; r; u, r'. let () := u in r'))
  %\end{verbatim}
  %\begin{align*}
  %  \mathit{lookup} &:
  %  \mathrm{WithList}_A \multimap \mathrm{List}_I \multimap \top \oplus A
  %  \\
  %                  &\coloneqq \lambda\mathit{xs}.~\mathrm{rec}\plr{xs; %
  %                    \mathrm{List}_I \multimap \top \oplus A; %
  %                    t.~\lambda i.~\mathrm{inl}\plr{%
  %                    \mathrm{foldr}\plr{i, \top \multimap \top; \lambda t'.~t'; %
  %                    u, r.~\textrm{let }{*} \coloneqq u\textrm{ in }r}\,t}; %
  %                    \mathit{xr}.~\lambda i.~%
  %                    \mathrm{foldr}\plr{i; %
  %                    A \with \plr{\top \oplus A} \multimap \top \oplus A; %
  %                    \lambda
  %                    }\,\mathit{xr}}
  %\end{align*}
\end{example}

As suggested by the notation, $\Box^{0{+}}$ may not be particularly canonical in
terms of what operations the usage context is stable under.
For example, we could also ask for $\forall \gr r.~\grQ \leq \gr r\grQ$, i.e.\
that the usage context is stable under any scaling.
This would allow us to include $\oc\gr r(-)$ in our grammar of strictly positive
functors.
In certain semirings, this could give us strange types, like, taking the
monotonicity semiring, the type
$\mu X.~I \oplus (\iota \otimes \oc\gr{\downarrow\downarrow} X)$,
which is the type of lists over the base type $\iota$ where alternating elements
are treated contravariantly and covariantly.
This may motivate us to restrict strictly positive functors to allow only a
subset of usage annotations.
Separately, in some contexts we may only need to know that $\grQ \leq \gr0$.
For example, in the linearity semiring, being less than or equal to $\gr0$
happens to imply closure under addition, because the only annotations less than
or equal to $\gr0$ are $\gr0$ and $\gr\omega$.
Also, in semirings for relevant systems, all annotations are stable under
addition (up to $\leq$ in the appropriate direction), so we need only check for
being less than or equal to $\gr0$.

Similar modalities named $\Box$ appear in the work of \citet{BB18} when they
deal with \emph{persistent predicates} in bunched logic, and in the work of
\citet{CK20} in a capability-aware calculus.
Both of these papers, in common with this thesis, aim to pick out the
\emph{pure} or \emph{safe} objects, i.e.\ those not depending on any external
resources.