### 2023-02-01 tex/rensub.tex

\chapter{Renaming and substitution for $\name$}\label{sec:ren-sub-lr}

In \cref{sec:semirings}, I defined my calculus of interest $\name$.
In this chapter, I develop the necessary syntactic metatheory for
specifying and implementing the substitution operation.
I follow the approach of \cref{sec:kits} using syntactic kits, but have to make
significant changes to the underlying notion of \emph{environment} before doing
so.
I give and informally motivate these changes to environments in
\cref{sec:lrkits}, and prove some properties of the new definition in
\cref{sec:lenv}.
Finally, I apply these new environments to the syntax of $\name$ in
\cref{sec:lrsub} to derive renaming and substitution operators.

\section{What are linear renaming and substitution?}\label{sec:lrkits}
\input{lrkits.tex}
\section{Properties of linear environments}\label{sec:lenv}
\input{lenv.tex}
\section{Substitution is admissible in \name{}}\label{sec:lrsub}
\input{lrsub.tex}
\section{Comparison with Petricek's substitution lemma}\label{sec:petricek}
\input{petricek.tex}
\section{Conclusion}\label{sec:ren-sub-lr-conc}

In this and the preceding chapter, I have developed a discipline for specifying
the syntax of linear and modal type systems, and furthermore developing the
syntactic metatheory of those type systems.
All of these are based on semirings, and the linear algebra arising from
considering a usage context full of semiring elements as a vector.

These developments can be seen in retrospect as a generalisation of the methods
explained in \cref{sec:simple}.
In terms of premise connectives in the syntactic rules, we have generalised from
just $\{\dot1, \dottimes\}$ to
$\{\dot1, \dottimes, I^*, \sep, \gr r\cdot{}, \Box^{0{+}}\}$, maintaining our
ability to keep the context implicit.
Similarly to how rule premises can require separation of usage annotations, our
new environments can require such a separation between their entries thanks to
the linear map they now contain.
I have generalised the key property of a kit from arbitrary weakening to
weakening by $\gr0$-annotated variables, and using that have produced a
substitution operation based on the same principles as that from
\cref{sec:kits}.

Having generalised all of the components --- namely the contexts, the syntax,
and the notion of environment --- the type of the substitution operation looks
the same as it did for intuitionistic ST$\lambda$C\@.
Being able to maintain this uniformity is a key step towards generalising the
rest of \cref{sec:simple} (i.e., \cref{sec:gen-sem,sec:gen-syn}), as I do in
\cref{sec:framework}.

\begin{itemize}
  \item Everything presented in this chapter is perfect.
  \item But equivalence of substitutions is hard to define/work with.
  \item The process of this development gives a template for defining
    simultaneous substitution for other substructural calculi.
\end{itemize}

I have made few unforced design decisions in this chapter, so there are few
aspects I can see future work improving upon.
However, future work may want to extend the work of this chapter, in which case
there are some unanswered questions.
Principle among these, in my mind, is dealing with equivalence/equality of
environments.
Intuitionistic environments $\rho, \sigma : \Gamma \env\V \Delta$ are equal
if and only if, for each type $A$ and each variable $x : \Delta \ni A$, we have
$\rho\,x = \sigma\,x$.
This follows from what we expect of equality of functions (function
extensionality).
Usage-aware environments, on the other hand, are $\Sigma$-types --- a way of
dividing up the usage annotations of $\Gamma$, and then a function producing
$\V$-values whose usage annotations come from that division.
Equality of $\Sigma$-types is tricky --- we need to equate the first components,
then rewrite the types of the second components by this equation before equating
them.
In practice, the equations rewriting other other equations build up so much that
I have given up on a first effort to give a treatment of the equational theory
of substitutions.
Note that recursively defined environments (\cref{def:lr-rec-env}) are also
$\Sigma$-types in cases where $\Delta$ is a concatenation of contexts, so that
definition does not clearly help.