### 2023-02-20 tex/mech-related.tex

There is a vast literature on formalisations of syntaxes with binding, which I
cannot possibly do justice to in a reasonably sized thesis chapter.
Instead, I limit myself to comparisons of the \citet{AACMM21} method I follow in
this thesis to just its closest related work.

\subsection{Autosubst}

\citet{Autosubst15} present the system \emph{Autosubst}, which provides various
tools for working with syntaxes with binding in the Coq proof assistant.
Autosubst is based on similar ideas to those \citeauthor{AACMM21} use:
de Bruijn-indexed terms with a distinguished variable rule and notion of
binding, acted upon by simultaneous renaming and substitution.

The simplest differences are essentially matters of choosing the encoding that
best fits the proof assistant being used.
Coq users tend to prefer using unindexed types and propositions indexed over
them --- in this case, a type of unscoped and untyped terms plus a
``well typed'' predicate --- whereas Agda users prefer to work with only well
formed data (well scoped and well typed terms).
The latter approach more readily allows us to show generically that substitution
preserves scoping and typing, but the former approach, conversely, allows for
bespoke proofs of such facts.
For example, one theorem of \citet{Autosubst15} is type preservation for
$\mathrm{CC}_\omega$, a dependent type system we cannot express using the
machinery of \citet{AACMM21}.
In principle, one could use \citeauthor{AACMM21}'s machinery as the basis of a
similar bespoke proof, but as far as I am aware, this has not been tried.

Another main difference is that Autosubst is presented to the user largely as
a black-box implementation of substitution and related lemmas, in contrast to
\citeauthor{AACMM21}'s work exposing the \AgdaRecord{Semantics} bundle to the
user, and having substitution be just one instance.
\citet{ACMM17} and \citet{AACMM21} provide many examples of traversals over
syntax using the same generic environment management as used by substitution.
However, the focus on substitution in Autosubst has meant that reasoning about
substitutions has been given more developed support.
For example, the library provides a tactic \texttt{autosubst} which automates
many equational proofs involving substitutions based on the $\sigma$-calculus of
\citet{ACCL91}.

An interesting feature of Autosubst is \emph{heterogeneous substitution}.
The motivation for heterogeneous substitution is to handle systems like system
F, where types and terms are syntactically distinct, but both feature binding
and require a substitution operation.
Furthermore, binding and substitution of types also affects the syntax of terms,
thanks to $\Lambda$ terms.
\citeauthor{AACMM21} provide no direct equivalent to heterogeneous substitution,
and it is unclear how well their work can handle polymorphic calculi.

\citet{Autosubst18} propose some modifications to Autosubst which, as far as I
can tell, have not yet been incorporated, but are presented in mechanised form
for the paper.
Some of these modifications aim to bring Autosubst into line with
\citeauthor{AACMM21}'s work, in particular taking semantic traversals as a
theoretical basis.
TODO: more.

\subsection{Fiore et al.}

\todo{This section is currently a random selection of paragraphs.}
Marcelo Fiore and various collaborators have a long line of work aiming for a
categorical account of variable-binding~\citep{FPT99,Fiore08,FH13,FH10,FM10}.
A recent, particularly relevant paper from this line of work is \citet{FS22},
which mechanises some of this work to obtain a framework similar in scope to
the work of \citet{AACMM21}.

Perhaps the main methodological difference between the works by
\citet{AACMM21} and \citet{FS22} is that \citet{AACMM21} start from typed
syntaxes and then give them semantics, whereas \citet{FS22} start by
characterising the category of models of a theory, and then later showing that
syntaxes form the initial model.
The latter approach means that \citeauthor{FS22} can do a significant amount of
theory not mentioning syntax in a compositional and structured style.
In contrast, the former approach is more lightweight when one wants to write
programs without a categorical specification.

I chose to base the work of this thesis on that of \citet{AACMM21} for largely
circumstantial reasons.
I only became aware of the work of \citet{FS22} when it was published, by which
time I had already completed the bulk of the work presented in this thesis.
Also, one of the authors of the former paper is my PhD supervisor, and the other
authors too were working nearby, so it was easy to discuss their work quickly
and informally.

In terms of the underlying theory, both works extend multi-sorted universal
algebra\todo{cite} with variable-binding.
However, the two extensions are subtly different.
Universal algebra already has a notion of variable, which supports renaming and
substitution, and which allows a given term to be evaluated when the free
variables of that term are assigned semantic values.
\citeauthor{AACMM21} reuse this notion of variable, and allow binding of such
variables in terms.
On the other hand, \citeauthor{FS22} recast the existing variables as
\emph{metavariables}, and introduce a new notion of (bindable) variables
separately into the syntax.
The resulting metavariables can then stand for arbitrary \emph{open} terms, and
thus each one remembers its context and requires a substitution whenever it is
used.

%Compared to the work of \citet{AACMM21}, which is what I chose to guide this
%chapter and the rest of this thesis, \citet{FS22} can be understood as providing
%many extra features.
%I happened not to be interested in these specific features, so have not
%implemented them, but future development of the work I present in this thesis
%could benefit from such features.
%The extra features include an externally compiled language of syntax and theory
%descriptions, internal support for metavariables, and a treatment of the
%category of models (of which terms are the initial model and what I call
%\AgdaFunction{sem} forms the universal morphism from terms to any other model).

\Citeauthor{FS22} use metavariables to form descriptions of relations over
terms, like an equational theory over the simply typed $\lambda$-calculus.
Terms have, as well as their context of variables, a context of metavariables,
and terms can contain a metavariable wherever they could contain a subterm.
There is then an operation of metavariable substitution, which substitutes terms
in the place of metavariables.
Metavariable substitution is used to instantiate the rules of described
relations/theories.
In contrast, \citeauthor{AACMM21} make do with the variables of the metalanguage
(i.e.\ Agda variables).
\todo{Congruence once and for all.}

Technologically, \citeauthor{FS22} provide an external domain-specific language
for syntax descriptions.
From such a description, a Python program generates some boilerplate Agda code
providing the types, the algebraic signature, the well typed terms, and a proof
that the terms are the initial model.
Using code generation like this resembles parts of the Autosubst plugin, as
opposed to the purely internal, generic programming-based solution given by
\citeauthor{AACMM21} and in this thesis.

\Citet[p.\ 19]{FS22} avoid sized types by defining their equivalent of the
functorial mapping \AgdaFunction{map-s} specialised to \AgdaFunction{sem}, with
these two functions being mutually recursive.
This is a standard solution, where code reuse is traded away in favour of
satisfying the termination checker and avoiding sized types.