### 2023-05-02 tex/linear.tex
[ writing ] mostly C3 related work

\chapter{Linearity and modality}\label{sec:linearity}

\begin{itemize}
  \item Motivation via prior work:
    \begin{itemize}
      \item Rust
      \item Session types
      \item ATS
      \item Complexity \citep{Hofmann03}
      \item Modal\ldots
    \end{itemize}
  \item Can \citet{AACMM21} capture linearity/modality?
    It's unclear based on standard presentations.
    Let's find one that looks good (DILL/PD), and see whether it works.
  \item In-depth introduction of DILL and PD
  \item It's close to working, but doesn't.
  \item Mention \citet{BBdePH93}, problems with substitution.
\end{itemize}

The techniques of \cref{sec:simple} were developed to handle idealised core
calculi, like the simply typed $\lambda$-calculus and its many variants.
Such core calculi are fairly close to some real intermediate languages,
\todo{cite} though many intermediate languages also have polymorphism in their
type systems, which I have not discussed, but has been considered by much prior
work\todo{cite}.

%Linear logic was introduced by \citet{girard87linear} as a tool of proof theory.
%Its distinguishing feature is that hypotheses, by default, must be used exactly
%once.
%I devote a chapter to introducing linear logic because it is better understood
%than any individual or general semiring-annotated calculi, and is thus an
%important point of comparison.
%Linear logic will be an instance of the semiring-annotated systems we study
%later.

\section{Intuitionistic linear logic}\label{sec:ill}
\input{ill.tex}

\section{DILL}\label{sec:dill}
\input{dill.tex}

\section{Mechanisation survey}
\input{linmech.tex}