### 2023-05-02 tex/simple.tex

\chapter{Mechanisation of simple types}\label{sec:simple}

In this chapter, I review and justify the family of approaches usually used to
represent simple type systems inside dependently typed proof assistants.
These approaches were first presented by \citet{AR99}, who showed a way of
representing well \emph{scoped} terms in a language with polymorphic recursion,
and extended the representation to well \emph{typed} terms in a language with
dependent types.
The representation of terms relies on indexing on both a context --- giving the
types of all the free variables --- and a type for the term itself.
A basic operation on terms is \emph{simultaneous substitution}, which replaces
each variable in the context by a term in another context.
This simultaneous substitution operation fits the form of composition in a
multicategory.
Specifically, in the case of a simple intuitionistic type system, this is the
composition of a Cartesian multicategory, which allows all of the terms being
substituted in to share a context.
\todo{Maybe delete multicategory remark.}

This chapter almost entirely reviews prior work.
In \cref{sec:terms}, I use a presentation of the dependently typed encoding of
the simply typed $\lambda$-calculus from \citet{AR99} to set notational
conventions.
\Cref{sec:kits} presents the unpublished work of \citet{McBride05},
successively deriving simultaneous renaming and simultaneous substitution for
the terms defined in \cref{sec:terms}.
The rest of the chapter generalises the shared core of renaming and substitution
in two dimensions: in \cref{sec:gen-sem} following \citet{ACMM17} to cover
semantic traversals, and in \cref{sec:gen-syn} following \citet{AACMM21} to
cover a whole range of simply typed syntaxes with binding, rather than just
a specific syntax.
Finally, I review some related work in \cref{sec:mech-related}.

\section{Term representation}\label{sec:terms}
\input{terms.tex}
\section{Renaming and substitution}\label{sec:kits}
\input{kits.tex}
\section{Generic semantics}\label{sec:gen-sem}
\input{gen-sem.tex}
\section{Generic syntax}\label{sec:gen-syn}
\input{gen-syn.tex}
%\section{Natural deduction vs sequent calculus}
%\input{ndsc.tex}
\section{Related work}\label{sec:mech-related}
\input{mech-related.tex}