### 2023-05-15 agda/PD.lagda.tex
[ writing ] about PD modal logic

\begin{code}
{-# OPTIONS --safe --without-K --postfix-projections #-}

module PD where

  data Ty : Set where
    ι : Ty
    □ : Ty → Ty

  data Jud : Set where
    true valid : Ty → Jud

  data True : Jud → Set where
    true : ∀ {A} → True (true A)
  data Valid : Jud → Set where
    valid : ∀ {A} → Valid (valid A)

  data TypeMatch (A : Ty) : Jud → Set where
    true : TypeMatch A (true A)
    valid : TypeMatch A (valid A)

  infix 50 _++_

  data Tree (X : Set) : Set where
    [_] : X → Tree X
    [] : Tree X
    _++_ : Tree X → Tree X → Tree X

  data All {X} (T : X → Set) : Tree X → Set where
    [_] : ∀ {x} → T x → All T [ x ]
    [] : All T []
    _++_ : ∀ {xt yt} → All T xt → All T yt → All T (xt ++ yt)

  data Any {X} (T : X → Set) : Tree X → Set where
    here : ∀ {x} → T x → Any T [ x ]
    ↙ : ∀ {xt yt} → Any T xt → Any T (xt ++ yt)
    ↘ : ∀ {xt yt} → Any T yt → Any T (xt ++ yt)

  map-Any : ∀ {X} {T U : X → Set} →
    (∀ {x} → T x → U x) → (∀ {xt} → Any T xt → Any U xt)
  map-Any f (here x) = here (f x)
  map-Any f (↙ i) = ↙ (map-Any f i)
  map-Any f (↘ i) = ↘ (map-Any f i)

  open import Data.Empty

  Any-⊥ : ∀ {X xt} → Any {X = X} (λ _ → ⊥) xt → ⊥
  Any-⊥ (↙ i) = Any-⊥ i
  Any-⊥ (↘ i) = Any-⊥ i

  open import Data.Product as ×
  open import Data.Sum as ⊎
  open import Function.Base using (id; _∘_; _|>_; case_of_)
  open import Relation.Binary.PropositionalEquality as ≡ using (_≡_; refl)

  lookup-All : ∀ {X} {T : X → Set} {xt x} → All T xt → Any (x ≡_) xt → T x
  lookup-All [ t ] (here refl) = t
  lookup-All (ttl ++ ttr) (↙ i) = lookup-All ttl i
  lookup-All (ttl ++ ttr) (↘ i) = lookup-All ttr i

  Ctx : Set
  Ctx = Tree Jud

  infix 40 _∋_ _⊢_

  _∋_ : Ctx → Ty → Set
  Γ ∋ A = Any (TypeMatch A) Γ

  view-∋-true-valid : ∀ {Γ A} → Γ ∋ A → Any (true A ≡_) Γ ⊎ Any (valid A ≡_) Γ
  view-∋-true-valid (here true) = inj₁ (here refl)
  view-∋-true-valid (here valid) = inj₂ (here refl)
  view-∋-true-valid (↙ v) = ⊎.map ↙ ↙ (view-∋-true-valid v)
  view-∋-true-valid (↘ v) = ⊎.map ↘ ↘ (view-∋-true-valid v)

  keep-valid : Ctx → Tree Ty
  keep-valid [ true A ] = []
  keep-valid [ valid A ] = [ A ]
  keep-valid [] = []
  keep-valid (Γ ++ Δ) = keep-valid Γ ++ keep-valid Δ

  all-valid : Tree Ty → Ctx
  all-valid [ A ] = [ valid A ]
  all-valid [] = []
  all-valid (As ++ Bs) = all-valid As ++ all-valid Bs

  filter-valid : Ctx → Ctx
  filter-valid = all-valid ∘ keep-valid

  pres-valid : ∀ {Γ A} →
    Any (valid A ≡_) Γ → Any (valid A ≡_) (filter-valid Γ)
  pres-valid (here refl) = here refl
  pres-valid (↙ i) = ↙ (pres-valid i)
  pres-valid (↘ i) = ↘ (pres-valid i)

  data _⊢_ (Γ : Ctx) : Ty → Set where
    var : ∀ {A} → Γ ∋ A → Γ ⊢ A
    □i : ∀ {A} → filter-valid Γ ⊢ A → Γ ⊢ □ A
    □e : ∀ {A B} → Γ ⊢ □ A → Γ ++ [ valid A ] ⊢ B → Γ ⊢ B

  record Env (T : Ctx → Ty → Set) (Γ Δ : Ctx) : Set where
    field
      true-part : ∀ {A} → Any (true A ≡_) Δ → T Γ A
      valid-part : ∀ {A} → Any (valid A ≡_) Δ → T (filter-valid Γ) A
  open Env public

  Ren = Env _∋_
  Sub = Env _⊢_

  view-Env-++ : ∀ {T Γ Δl Δr} → Env T Γ (Δl ++ Δr) → Env T Γ Δl × Env T Γ Δr
  view-Env-++ ρ .proj₁ .true-part = ρ .true-part ∘ ↙
  view-Env-++ ρ .proj₁ .valid-part = ρ .valid-part ∘ ↙
  view-Env-++ ρ .proj₂ .true-part = ρ .true-part ∘ ↘
  view-Env-++ ρ .proj₂ .valid-part = ρ .valid-part ∘ ↘

  filter-valid-Valid : ∀ {Γ} → All Valid (filter-valid Γ)
  filter-valid-Valid {Γ = [ true A ]} = []
  filter-valid-Valid {Γ = [ valid A ]} = [ valid ]
  filter-valid-Valid {Γ = []} = []
  filter-valid-Valid {Γ = Γ ++ Δ} = filter-valid-Valid ++ filter-valid-Valid

  -- Valid-in-filter-valid :
  --   ∀ {Γ A} → Any (valid A ≡_) Γ → Any (valid A ≡_) (filter-valid Γ)
  -- Valid-in-filter-valid i = {!!}

  filter-valid-⊆ : ∀ {T Γ} → Any T (filter-valid Γ) → Any T Γ
  filter-valid-⊆ {Γ = [ valid A ]} i = i
  filter-valid-⊆ {Γ = Γ ++ Δ} (↙ i) = ↙ (filter-valid-⊆ i)
  filter-valid-⊆ {Γ = Γ ++ Δ} (↘ i) = ↘ (filter-valid-⊆ i)

  filter-valid-idem : ∀ {Γ} → filter-valid (filter-valid Γ) ≡ filter-valid Γ
  filter-valid-idem {Γ = [ true A ]} = refl
  filter-valid-idem {Γ = [ valid A ]} = refl
  filter-valid-idem {Γ = []} = refl
  filter-valid-idem {Γ = Γ ++ Δ} =
    ≡.cong₂ _++_ filter-valid-idem filter-valid-idem

  filter-valid-Ren : ∀ {Γ} → Ren Γ (filter-valid Γ)
  filter-valid-Ren .true-part i = case (lookup-All filter-valid-Valid i) of λ ()
  filter-valid-Ren .valid-part i = map-Any (λ { refl → valid }) i

  module _ {T : Ctx → Ty → Set}
    (ren : ∀ {Γ Δ A} → Ren Γ Δ → T Δ A → T Γ A)
    where

    lookup : ∀ {Γ Δ A} → Env T Γ Δ → Δ ∋ A → T Γ A
    lookup ρ v = view-∋-true-valid v |>
      [ ρ .true-part , ren filter-valid-Ren ∘ ρ .valid-part ]′

  module BindEnv {T : Ctx → Ty → Set}
    (vr : ∀ {Γ A} → Any (TypeMatch A) Γ → T Γ A)
    (wk : ∀ {Γ A Θ} → T Γ A → T (Γ ++ Θ) A)
    where

    vt : ∀ {Γ A} → Any (true A ≡_) Γ → T Γ A
    vt = vr ∘ map-Any (λ { refl → true })
    vv : ∀ {Γ A} → Any (valid A ≡_) Γ → T Γ A
    vv = vr ∘ map-Any (λ { refl → valid })

    bindEnv : ∀ {Γ Δ Θ} → Env T Γ Δ → Env T (Γ ++ Θ) (Δ ++ Θ)
    bindEnv ρ .true-part (↙ i) = wk (ρ .true-part i)
    bindEnv ρ .true-part (↘ i) = vt (↘ i)
    bindEnv ρ .valid-part (↙ i) = wk (ρ .valid-part i)
    bindEnv ρ .valid-part (↘ i) = vv (↘ (pres-valid i))

  module _ {T : Ctx → Ty → Set} where

    Env-filter-valid : ∀ {Γ Δ} →
      Env T Γ Δ → Env T (filter-valid Γ) (filter-valid Δ)
    Env-filter-valid ρ .true-part i =
      case (lookup-All filter-valid-Valid i) of λ ()
    Env-filter-valid ρ .valid-part i =
      let t = ρ .valid-part (filter-valid-⊆ i) in
      ≡.subst (λ ∙ → T ∙ _) (≡.sym filter-valid-idem) t

  module Trav {T : Ctx → Ty → Set}
    (vr : ∀ {Γ A} → Any (TypeMatch A) Γ → T Γ A)
    (ren : ∀ {Γ Δ A} → Ren Γ Δ → T Δ A → T Γ A)
    (tm : ∀ {Γ A} → T Γ A → Γ ⊢ A)
    where

    wk : ∀ {Γ A Θ} → T Γ A → T (Γ ++ Θ) A
    wk = ren λ where
      .true-part i → ↙ (map-Any (λ { refl → true }) i)
      .valid-part i → ↙ (map-Any (λ { refl → valid }) (pres-valid i))

    open BindEnv vr wk

    trav : ∀ {Γ Δ A} → Env T Γ Δ → Δ ⊢ A → Γ ⊢ A
    trav ρ (var x) = tm (lookup ren ρ x)
    trav ρ (□i M) = □i (trav (Env-filter-valid ρ) M)
    trav ρ (□e M N) = □e (trav ρ M) (trav (bindEnv ρ) N)
\end{code}

\begin{code}
  ren^∋ : ∀ {Γ Δ A} → Env _∋_ Γ Δ → Δ ∋ A → Γ ∋ A
  ren^∋ ρ v = view-∋-true-valid v |>
    [ ρ .true-part , filter-valid-⊆ ∘ ρ .valid-part ]′

  ren^⊢ : ∀ {Γ Δ A} → Env _∋_ Γ Δ → Δ ⊢ A → Γ ⊢ A
  ren^⊢ = Trav.trav id ren^∋ var

  sub^⊢ : ∀ {Γ Δ A} → Env _⊢_ Γ Δ → Δ ⊢ A → Γ ⊢ A
  sub^⊢ = Trav.trav var ren^⊢ id
\end{code}