### 2023-05-15 tex/linear.tex

\chapter{Linearity and modality}\label{sec:linearity}

\begin{itemize}
  \item Motivation via prior work:
    \begin{itemize}
      \item Rust
      \item Session types
      \item ATS
      \item Complexity \citep{Hofmann03}
      \item Modal\ldots
    \end{itemize}
  \item Can \citet{AACMM21} capture linearity/modality?
    It's unclear based on standard presentations.
    Let's find one that looks good (DILL/PD), and see whether it works.
  \item In-depth introduction of DILL and PD
  \item It's close to working, but doesn't.
  \item Mention \citet{BBdePH93}, problems with substitution.
\end{itemize}

The techniques of \cref{sec:simple} were developed to handle idealised core
calculi, like the simply typed $\lambda$-calculus and its many variants.
Such core calculi are fairly close to some real intermediate languages,
\todo{cite} though many intermediate languages also have polymorphism in their
type systems, which I have not discussed, but has been considered by much prior
work~\citep{POPLmark}.

Another potential feature --- required for many desired semantics but only
relatively recently appearing in programming languages --- that is not handled
by the techniques of \cref{sec:simple} is restrictions on how variables can be
used.
In the simple type systems covered in \cref{sec:simple}, the only restriction on
the use of a variable is that it is used as an expression of the appropriate
type.
However, we could also consider having somehow ``guarded'' subexpressions in
which some variables bound outside cannot occur, or systems in which the use of
a variable in one part of an expression precludes its use in another part of
that expression (the variable having been ``used up'').
I will introduce and motivate some such systems in this chapter, and argue that
they are impossible to capture directly in the framework presented in
\cref{sec:gen-syn}.

Languages and systems implementing features inspired by linear logic include
Rust~\citep{MK14,Rust}, ATS~\citep{Xi04,ZX05}, and various implementations of
session types~\citep{HLVCCDMPRTVTZ16}.
Additionally, forms of linearity have been used in theoretical work, for example
to bound computational complexity~\citep{GSS92,Hofmann03} or maintain
differentiability~\citep{ER03,Ehrhard18}.
As for modality, \citet{IMO20} show that common data-flow analyses for
imperative languages can be recast as modal type systems, though they have
generally been developed in an ad hoc fashion.
Such a recasting may be useful to understand the metatheoretic properties of
programs passing such static analyses.
More user-facing implementations of modal type systems include the system for
stack allocation in OCaml~\citep{DW22} and the approach to erasure found in
Agda and Idris\todo{cite}.

In this chapter\ldots\todo{Chapter map}

\section{Intuitionistic S4 modal logic}\label{sec:modal}
\input{modal.tex}

\section{Intuitionistic linear logic}\label{sec:ill}
\input{ill.tex}

\section{DILL}\label{sec:dill}
\input{dill.tex}

\section{Mechanisation survey}
\input{linmech.tex}