### 2023-07-25 tex/ill.tex

Where modal logics can be seen as additions to an underlying classical or
intuitionistic logic, linear logic is a much more radical change.
Formally, many modal logics, including (I)S4, are conservative extensions of
the underlying classical or intuitionistic logic, meaning that statements not
mentioning the modal operators $\Box$ and $\Diamond$ are provable in modal logic
if and only if they are provable in the underlying non-modal logic.
In linear logic, however, we a priori severely restrict provability of basic
formulae, and then use modalities to recover the strength of classical or
intuitionistic logic.

In this thesis, I consider intuitionistic linear logic (ILL).
ILL can be understood as a logic of resources.
Whereas in classical logic we read a proposition $A$ to implicitly mean ``$A$ is
true'', we may read a proposition $A$ of linear logic to implicitly mean ``I
have an $A$''.
A sequent $A \vdash B$ can then be understood as saying ``if I have an $A$, I
can give it up so as to have a $B$'', and the corresponding implication
$A \multimap B$ can be understood as saying ``I have a \emph{mechanism} for
having a $B$ if I have and am willing to give up an $A$''.

The resource interpretation of intuitionistic linear logic is often explained by
analogy with a vending machine.
Suppose we have a vending machine which takes pound coins and sells bottles of
water for \pounds{}1 and chocolate bars for \pounds{}2.
We can represent the latter two facts by the judgements
$\textrm{\pounds}1 \vdash \mathrm{water}$ and
$\textrm{\pounds}1, \textrm{\pounds}1 \vdash \mathrm{bar}$, respectively.
Notice that having two \pounds{}1 coins is very different from having just one
\pounds{}1 coin.
We can represent having compound objects using $\otimes$-products
(tensor-products), so that having a \pounds{}1 coin together with a chocolate
bar is represented as $\textrm{\pounds}1\otimes\mathrm{bar}$, and the mechanism
for getting a chocolate bar from the vending machine is described by
$\plr{\textrm{\pounds}1 \otimes \textrm{\pounds}1} \multimap \mathrm{bar}$.
The entire protocol of one transaction with the vending machine is represented
as
$\plr{\textrm{\pounds}1 \multimap \mathrm{water}} \with
\plr{\plr{\textrm{\pounds}1 \otimes \textrm{\pounds}1} \multimap \mathrm{bar}}$,
where the $\with$-product (with-product) represents giving the user a choice of
which of the two conjuncts to have.
Finally, that we can do multiple transactions with the vending machine is
represented as
$\oc\plr{\plr{\textrm{\pounds}1 \multimap \mathrm{water}} \with
\plr{\plr{\textrm{\pounds}1 \otimes \textrm{\pounds}1} \multimap \mathrm{bar}}}
$, where the $\oc$-modality (read ``bang'' or ``of course'') represents having
an arbitrary number of copies of its argument.

Such reasoning is useful in computer science because it allows us to distinguish
what we had before some event with what we got after it.
In both intuitionistic and classical logic, if we had a \pounds{}1 coin and a
way to turn a \pounds{}1 coin into a bottle of water, we'd be able to have both
the \pounds{}1 coin and the bottle of water at once.
Instead, linear logic acknowledges that the \pounds{}1 coin was spent, so at one
time we had just the \pounds{}1 coin, and at a later time we had just the bottle
of water.
In real applications, such reasoning is useful in, for example, dealing with
mutable state~\citep{MK19} --- in which, when we mutate the state, we forget the
old value and have only the new value --- and producing a session type system~%
\citep{Wadler12} --- in which the protocol progresses as messages are sent.

In this section, I will carefully introduce intuitionistic linear logic so as to
make the intuitive readings more precise.

\subsection{The multiplicative-additive fragment}

The multiplicative-additive fragment of linear logic (MALL) is the fragment
where all hypotheses are linear (must be used exactly once).
I will extend MALL with the \emph{exponential} modality in
\cref{sec:bang-modality}.
MALL is unable to embed intuitionistic or classical logic, as MALL is unable to
reflect any of the discarding or duplication that can be done in proofs using
weakening or contraction.

In short, the syntax of intuitionistic MALL can be described as intuitionistic
logic with the structural rules of \emph{weakening} and \emph{contraction}
removed.
However, without the presence of weakening and contraction, we have to be more
careful about the rules we state, so as not to accidentally admit weakening
and contraction.
The lack of these structural rules also allows us to observe a new phenomenon:
the distinction (at the level of provability) between \emph{additive} and
\emph{multiplicative} formulations of existing connectives (in particular, the
conjunction connective).

I present MALL in \cref{fig:mall} in a sequent calculus style, as it was
presented by \citet{girard87linear}.

To encode what it means to use a hypothesis \emph{exactly once}, we first need
to decide what counts as a use.
The simplest case is that the identity sequent counts as a single use of its
sole hypothesis, and conversely does not count as a use of any other hypotheses.
For sequential proofs, created by the \TirName{Cut} rule, if we have a proof of
$A$ using $\Gamma$, and a proof of $B$ using $\Delta$ and $A$, then we have a
proof of $B$ transitively using $\Delta$ and $\Gamma$.
The exchange rule \TirName{Exch} says that use is invariant under permutation.

For the logical connectives, we have genuine choices as to what it means to use
them.
Two cases --- disjunction ($\oplus$) and (linear) implication ($\multimap$) ---
are somewhat intuitive from intuitionistic logic.
A canonical proof of a disjunction is a tag and a proof of one of the two
disjuncts.
This suggests that a proof of a disjunction only uses the same hypotheses as
the proof of the disjunct we actually choose, with the other disjunct being
irrelevant.
Correspondingly, when we use a disjunction hypothesis, we will only actually use
one of the cases, so each branch should use the same hypotheses.
For implication, use is sequential like with the \TirName{Cut} rule, and its
left rule is more or less the only choice that allows use of the surrounding
hypotheses.

For conjunction, there are two choices: Either the conjuncts \emph{together} use
all of the hypotheses, or each of the conjuncts \emph{individually} uses all of
the hypotheses.
The former choice gives us the tensor-product ($\otimes$), and the latter choice
gives us the with-product ($\with$).
These products are equivalent up to provability in logics with weakening and
contraction, but distinct in linear logic.

\begin{figure}
  \begin{align*}
    A, B, C &\Coloneqq X \mid I \mid A \otimes B \mid A \multimap B
              \mid 0 \mid A \oplus B \mid \top \mid A \with B \\
    \Gamma, \Delta, \Theta &\Coloneqq {\cdot} \mid \Gamma, A
  \end{align*}
  \begin{mathpar}
    \ebrule{%
      \infer0[Id]{A \vdash A}
    }

    \and

    \ebrule{%
      \hypo{\Gamma \vdash A}
      \hypo{\Delta, A \vdash B}
      \infer2[Cut]{\Gamma, \Delta \vdash B}
    }

    \and

    \ebrule{%
      \hypo{\Gamma, B, A, \Delta \vdash C}
      \infer1[Exch]{\Gamma, A, B, \Delta \vdash C}
    }

    \and

    \ebrule{%
      \hypo{\Gamma \vdash C}
      \infer1[$I$-L]{\Gamma, I \vdash C}
    }

    \and

    \ebrule{%
      \infer0[$I$-R]{{\cdot} \vdash I}
    }

    \and

    \ebrule{%
      \hypo{\Gamma, A, B \vdash C}
      \infer1[$\otimes$-L]{\Gamma, A \otimes B \vdash C}
    }

    \and

    \ebrule{%
      \hypo{\Gamma \vdash A}
      \hypo{\Delta \vdash B}
      \infer2[$\otimes$-R]{\Gamma, \Delta \vdash A \otimes B}
    }

    \and

    \ebrule{%
      \hypo{\Gamma \vdash A}
      \hypo{\Delta, B \vdash C}
      \infer2[$\multimap$-L]{\Gamma, \Delta, A \multimap B \vdash C}
    }

    \and

    \ebrule{%
      \hypo{\Gamma, A \vdash B}
      \infer1[$\multimap$-R]{\Gamma \vdash A \multimap B}
    }

    \and

    \ebrule{%
      \infer0[$0$-L]{\Gamma, 0 \vdash C}
    }

    \and

    \text{(no $0$-R)}

    \and

    \ebrule{%
      \hypo{\Gamma, A \vdash C}
      \hypo{\Gamma, B \vdash C}
      \infer2[$\oplus$-L]{\Gamma, A \oplus B \vdash C}
    }

    \and

    \ebrule{%
      \hypo{\Gamma \vdash A_i}
      \infer1[$\oplus$-R$_i$]{\Gamma \vdash A_0 \oplus A_1}
    }

    \and

    \text{(no $\top$-L)}

    \and

    \ebrule{%
      \infer0[$\top$-R]{\Gamma \vdash \top}
    }

    \and

    \ebrule{%
      \hypo{\Gamma, A_i \vdash C}
      \infer1[$\with$-L$_i$]{\Gamma, A_0 \with A_1 \vdash C}
    }

    \and

    \ebrule{%
      \hypo{\Gamma \vdash A}
      \hypo{\Gamma \vdash B}
      \infer2[$\with$-R]{\Gamma \vdash A \with B}
    }
  \end{mathpar}
  \caption{Multiplicative-additive fragment of linear logic}
  \label{fig:mall}
\end{figure}

Implication ($\multimap$) and the tensor-product ($\otimes$, $I$) comprise the
\emph{multiplicative} fragment, while disjunction ($\oplus$, $0$) and the
with-product ($\with$, $\top$) comprise the \emph{additive} fragment.
Categorically, the additive fragment corresponds to products and coproducts,
while the multiplicative fragment corresponds to multicategorical tensor
products and closure.

\subsection{The $\oc$-modality}\label{sec:bang-modality}

\begin{figure}
  \begin{align*}
    A, B, C &\Coloneqq \dots \mid \oc A
  \end{align*}
  \begin{mathpar}
    \ebrule{%
      \hypo{\oc\Gamma \vdash A}
      \infer1[Promotion]{\oc\Gamma \vdash \oc A}
    }

    \and

    \ebrule{%
      \hypo{\Gamma, A \vdash B}
      \infer1[Dereliction]{\Gamma, \oc A \vdash B}
    }

    \and

    \ebrule{%
      \hypo{\Gamma \vdash B}
      \infer1[Weakening]{\Gamma, \oc A \vdash B}
    }

    \and

    \ebrule{%
      \hypo{\Gamma, \oc A, \oc A \vdash B}
      \infer1[Contraction]{\Gamma, \oc A \vdash B}
    }
  \end{mathpar}
  \caption{The sequent calculus rules for the $\oc$-modality}
  \label{fig:bang-seq}
\end{figure}

\Cref{fig:bang-seq} shows the rules we can add to MALL to get the full sequent
calculus for intuitionistic linear logic (ILL).
In ILL, $\oc A$ is defined
to be a proposition whose occurrences as antecedents can be deleted
(\TirName{Weakening}) and duplicated (\TirName{Contraction}), from which we can
extract $A$ (\TirName{Dereliction}), and which we can form from a conclusion
$A$ only when all antecedents are of the form $\oc X$ for some proposition $X$
(\TirName{Promotion}).
In short, $\oc A$ can be seen as an intuitionistic version of $A$, supporting
all of the structural rules of LJ, and only being able to be formed when it
does not depend on anything linear.

The \TirName{Dereliction} rule is comparable to the \TirName{T} axiom we saw in
\cref{sec:modal}.
The \TirName{Promotion} rule is exactly like the \TirName{Promotion} rule we saw
in \cref{sec:modal}, and for the same reasons we want to avoid having such a
rule in a natural deduction system for intuitionistic linear logic.
Note that \TirName{Promotion} is not problematic in the sequent calculus, at
least insofar as it maintains \TirName{Cut}-elimination.

Additional to the problems with \TirName{Promotion} in a natural deduction
system, defining the $\oc$-modality as in \cref{fig:bang-seq} has the odd
feature that, unlike all of the other connectives of ILL, $\oc$ is not
characterised by a universal property.
This can be seen by the fact that taking the rules for $\oc$ and replacing each
occurrence of $\oc$ by a fresh connective $\oc'$ produces a logically distinct
connective.
One cannot produce any derivation of $\oc' A \vdash \oc A$ because
\TirName{Promotion} does not apply when there are antecedents not of the form
$\oc X$.
This lack of characterisation also holds of the rule and axioms we first gave to
the $\Box$-modality in \cref{fig:S4-axioms}, and essentially any presentation
that does not incorporate the modality into the judgemental structure of the
calculus.

A noteworthy solution to the substitution problem of \TirName{Promotion} is
given by \citet{BBdePH93}.
The solution they give is compatible with the form of sequents introduced above,
but works purely in terms of right-rules like standard natural deduction
calculi.
I give their rules for the $\oc$-modality in \cref{fig:BBdePH}.
The \TirName{Weakening} and \TirName{Contraction} rules are reminiscent of
pattern-matching elimination principles, allowing us to form an inhabitant of
the type of interest ($\oc A$) and then continue producing the originally
desired result ($B$) in an updated context.
However, somewhat unusually, we can choose which of the two ``pattern-matching''
principles to use, depending on whether we want to delete or duplicate the
value of type $\oc A$.
Another elimination-like rule is \TirName{Dereliction}, which straightforwardly
lets us derive $A$ from $\oc A$ similarly to the corresponding sequent calculus
rule from \cref{fig:bang-seq}, but on the right.

\begin{figure}
  \begin{mathpar}
    \ebrule{%
      \hypo{\Delta_1 \vdash \oc A_1}
      \hypo{\cdots}
      \hypo{\Delta_n \vdash \oc A_n}
      \hypo{\oc A_1, \ldots, \oc A_n \vdash B}
      \infer4[Promotion]{\Delta_1, \ldots, \Delta_n \vdash \oc B}
    }
    \and
    \ebrule{%
      \hypo{\Gamma \vdash \oc A}
      \infer1[Dereliction]{\Gamma \vdash A}
    }
    \and
    \ebrule{%
      \hypo{\Gamma \vdash \oc A}
      \hypo{\Delta \vdash B}
      \infer2[Weakening]{\Gamma, \Delta \vdash B}
    }
    \and
    \ebrule{%
      \hypo{\Gamma \vdash \oc A}
      \hypo{\Delta, \oc A, \oc A \vdash B}
      \infer2[Contraction]{\Gamma, \Delta \vdash B}
    }
  \end{mathpar}
  \caption{The \citet{BBdePH93} rules for the $\oc$-modality}
  \label{fig:BBdePH}
\end{figure}

The \citeauthor{BBdePH93} \TirName{Promotion} rule is more formally complex, but
can be understood as follows.
We notice that the problem with the \TirName{Promotion} rule of
\TirName{fig:bang-seq} when added to a natural deduction system is that it
restricts the types of assumptions in the concluding sequent.
In other words, the sequent calculus \TirName{Promotion} rule is acting like
both a right-rule and, problematically, a left-rule, to some degree.
Acting on the context in a natural deduction system risks making substitution
inadmissible, as happens in this case.
Instead, \citeauthor{BBdePH93} paraphrase this restriction on the context as the
construction of a \emph{new} context $\oc A_1, \ldots, \oc A_n$ for the primary
subderivation.
Heuristically, this new \TirName{Promotion} rule admits substitution because any
substitution will pass into the relevant premises deriving $\oc A_1$ to
$\oc A_n$.
More abstractly, we can notice that the first $n$ premises essentially form an
explicit substitution from $\Delta_1, \ldots, \Delta_n$ to
$\oc A_1, \ldots, \oc A_n$, so any further substitution applied
to a \TirName{Promotion}-headed term is precomposed onto this explicit
substitution~\citep{ACCL91}.

The \citeauthor{BBdePH93} \TirName{Promotion} rule is a clever solution to the
substitution problem of (intuitionistic) linear logic, but the resulting system
comes with its own problems.
If we were to write programs in a language directly implementing the
\citeauthor{BBdePH93} calculus, then the \TirName{Promotion} rule would be a
pain point because it makes us rebind all of the $\oc$-typed variables we need to
new variables.
Additionally, we have to be explicit everywhere about weakening and contraction,
potentially making working with $\oc$-typed variables much more fiddly than the
corresponding variables would be in a non-substructural programming language.
These syntactic annoyances may be worked around in a realistic implementation by
some new elaboration procedures, but we would prefer not to rely on such
elaboration if there is an acceptable core calculus which more closely matches
the programs we want to write.
Therefore, I investigate such a system, Dual Intuitionistic Linear Logic, in the
following subsection.

\subsection{Dual Intuitionistic Linear Logic}\label{sec:dill}
\input{dill.tex}